---
title: "02_methods_descriptives"
author: "Tim"
date: "2026-01-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(janitor)
library(rlang)
library(scales)# for percent formatting
library(here) # robust paths within the repo
library(stringr)
```

```{r}
path_in <- here("data", "ica_methods.csv")

dir.create(here("outputs"), showWarnings = FALSE)

raw <- readr::read_csv(
file = path_in,
col_types = readr::cols(.default = readr::col_character()),
na = c("", "NA", "N/A", "na", "n/a", "null", "NULL")
)

dat <- raw %>% clean_names()

glimpse(dat) #sanity check
```

```{r}

binary_cols <- c(
  "matrix",
  "short_scales",
  "scale_search",
  "prisma_a",
  "prisma",
  "inclusion_criteria_applicable",
  "inclusion_binary",
  "clinical_relevance",
  "language_mentioned",
  "availability_mentioned",
  "release_date_mentioned",
  "frequency",
  "min_scale_citation",
  "exclusion_criteria",
  "multiple_analyses",
  "within_scale_item_collapse",
  "compound_items",
  "idiosyncratic_items",
  "reverse_worded_items",
  "coding_team_reported",
  "blinded",
  "codebook",
  "software_mentioned",
  "item_exclusions",
  "irr_reported",
  "preregistered_rules",
  "coding_scheme_independent_mentioned",
  "mean_overlap",
  "min_overlap",
  "max_overlap",
  "length_correlation_noted"
)

# Select only binary columns
dat_binary_raw <- dat %>%
  select(study_id, any_of(binary_cols))
``` 

```{r}
to_logical_strict <- function(x, colname) {
  # Leave logical as-is
  if (is.logical(x)) return(x)

  x0 <- as.character(x)
  x1 <- tolower(trimws(x0))

  true_vals  <- c("true", "t", "yes", "y", "1")
  false_vals <- c("false", "f", "no", "n", "0")

  out <- dplyr::case_when(
    is.na(x1) ~ NA,
    x1 %in% true_vals  ~ TRUE,
    x1 %in% false_vals ~ FALSE,
    TRUE ~ NA
  )

  # In case of typos: any non-NA value not in the allowed encodings
  bad <- which(!is.na(x1) & !(x1 %in% c(true_vals, false_vals)))
  if (length(bad) > 0) {
    bad_vals <- sort(unique(x0[bad]))
    rlang::abort(paste0(
      "Non-boolean value(s) detected in column '", colname, "': ",
      paste(bad_vals, collapse = ", "),
      ". Example row indices: ", paste(head(bad, 10), collapse = ", ")
    ))
  }

  out
}

dat_binary <- dat_binary_raw %>%
  mutate(across(all_of(binary_cols), ~ to_logical_strict(.x, cur_column())))

```

```{r}
binary_counts <- dat_binary %>%
  pivot_longer(  #long table with one row per (paper × variable)
    cols = all_of(binary_cols),
    names_to = "variable",
    values_to = "value"
  ) %>%
  mutate(value = case_when( #converts logicals into `"TRUE"/"FALSE"/"NA"` labels
    is.na(value) ~ "NA",
    value ~ "TRUE",
    !value ~ "FALSE"
  )) %>%
  count(variable, value, name = "n") %>%  #counts how many of each label occurs per variable
  group_by(variable) %>%
  mutate(n_total = sum(n)) %>% #computes the total number of observations per variable while still in long format
  ungroup() %>%
  pivot_wider(  #spreads the counts into columns 'TRUE', 'FALSE' and 'NA' and keeps the n_total
    names_from = value,
    values_from = n,
    values_fill = 0
  ) %>%
  arrange(desc(TRUE))

binary_counts
```

```{r}
binary_overall <- dat_binary %>%
  summarise(  #computes per-column counts
    across(
      all_of(binary_cols),
      list(
        n_true  = ~ sum(.x %in% TRUE, na.rm = TRUE),
        n_false = ~ sum(.x %in% FALSE, na.rm = TRUE),
        n_na    = ~ sum(is.na(.x)),
        n_coded = ~ sum(!is.na(.x))
      ),
      .names = "{.col}__{.fn}"  #attempt to give systematic names like "irr_reported_n_true"
    )
  ) %>%
  pivot_longer(  #splits the names into 'variable" (the original column name) and '.value' (which statistic the number belongs to)
    cols = everything(),
    names_to = c("variable", ".value"),
    names_sep = "__"
  ) %>%
  mutate(
    prop_true = n_true / n_coded,  #proportion TRUE among all coded cases only
    percent_true = scales::percent(prop_true, accuracy = 0.1),
    report = paste0(n_true, "/", n_coded, " (", percent_true, ")")  #manuscript-friendly string
  ) %>%
  arrange(desc(prop_true))

binary_overall %>% select(variable, n_true, n_coded, n_na, report)

readr::write_csv(binary_overall, here("outputs", "binary_overall.csv"))

```

# Step 4: Conditional reporting rates for gated variables
There are a couple of key "gates" in the binary dataset: 
1. `prisma` only matters when `prisma_a == TRUE`
2. “selection criteria” only matter when `scale_search == TRUE`

This is done as: **filter to applicable subset → summarise the target variables**

## Step 4a: PRISMA reporting among applicable studies (`prisma_a == TRUE`)
### Why this step?
It enables to report a conditional statement and avoids penalising studies where PRISMA is structurally not applicable

```{r}
prisma_cond <- dat_binary %>%
  filter(prisma_a == TRUE) %>%
  summarise(
    n_applicable = n(),
    n_true  = sum(prisma %in% TRUE, na.rm = TRUE),
    n_false = sum(prisma %in% FALSE, na.rm = TRUE),
    n_na    = sum(is.na(prisma)),
    n_coded = sum(!is.na(prisma))
  ) %>%
  mutate(
    prop_true = n_true / n_coded,
    percent_true = scales::percent(prop_true, accuracy = 0.1),
    report = paste0(n_true, "/", n_coded, " (", percent_true, ")")
  )

prisma_cond

```


## Step 4b: Selection-criteria reporting among manual searches (`scale_search == TRUE`)

```{r}
criteria_cols <- c(
  "inclusion_binary",
  "clinical_relevance",
  "language_mentioned",
  "availability_mentioned",
  "release_date_mentioned",
  "frequency",
  "min_scale_citation",
  "exclusion_criteria"
)

criteria_cond <- dat_binary %>%
  filter(scale_search == TRUE) %>%
  summarise(
    across(
      all_of(criteria_cols),
      list(
        n_true  = ~ sum(.x %in% TRUE, na.rm = TRUE),
        n_false = ~ sum(.x %in% FALSE, na.rm = TRUE),
        n_na    = ~ sum(is.na(.x)),
        n_coded = ~ sum(!is.na(.x))
      ),
      .names = "{.col}__{.fn}"
    )
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = c("variable", ".value"),
    names_sep = "__"
  ) %>%
  mutate(
    prop_true = n_true / n_coded,
    percent_true = scales::percent(prop_true, accuracy = 0.1),
    report = paste0(n_true, "/", n_coded, " (", percent_true, ")"),
    condition = "Among studies with manual scale search (scale_search == TRUE)"
  ) %>%
  arrange(desc(prop_true))

criteria_cond %>% select(condition, variable, n_true, n_coded, n_na, report)

```

```{r}
readr::write_csv(prisma_cond, here("outputs", "prisma_conditional.csv"))
readr::write_csv(criteria_cond, here("outputs", "criteria_conditional.csv"))

```

#Step 5: Descriptives for "homogenous" numeric columns

```{r}
# "clean"/homogenous numeric columns 
numeric_cols_clean <- c(
  "n_scales",
  "initial_coding_team_size",
  "n_rater",
  "mean_overlap_nr",
  "min_overlap_nr",
  "max_overlap_nr"
)

numeric_cols_clean_present <- intersect(numeric_cols_clean, names(dat))

dat_num <- dat %>%
  mutate(across(all_of(numeric_cols_clean_present), readr::parse_number))

# NA rates after parsing (quick QA)
numeric_na_rates <- dat_num %>%
  summarise(across(all_of(numeric_cols_clean_present),
                   ~ mean(is.na(.x)))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "prop_na") %>%
  mutate(percent_na = scales::percent(prop_na, accuracy = 0.1)) %>%
  arrange(desc(prop_na))

numeric_na_rates

``` 

```{r}
summarise_numeric <- function(x) {
  x <- x[!is.na(x)]
  if (length(x) == 0) {
    return(tibble(
      n = 0, mean = NA_real_, sd = NA_real_, median = NA_real_, iqr = NA_real_,
      min = NA_real_, max = NA_real_, q25 = NA_real_, q75 = NA_real_,
      n_outliers_iqr = NA_integer_
    ))
  }

  q <- quantile(x, probs = c(.25, .75), names = FALSE)
  iqr_val <- q[2] - q[1]
  lower <- q[1] - 1.5 * iqr_val
  upper <- q[2] + 1.5 * iqr_val
  n_out <- sum(x < lower | x > upper)

  tibble(
    n = length(x),
    mean = mean(x),
    sd = sd(x),
    median = median(x),
    iqr = IQR(x),
    min = min(x),
    max = max(x),
    q25 = q[1],
    q75 = q[2],
    n_outliers_iqr = n_out
  )
}

numeric_overall <- dat_num %>%
  summarise(across(all_of(numeric_cols_clean_present), summarise_numeric)) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "stats") %>%
  unnest_wider(stats)

numeric_overall

```

```{r}
readr::write_csv(numeric_overall, here("outputs", "numeric_overall.csv"))

```

```{r}
dat_n_scales_num <- dat %>%
  mutate(n_scales_num = parse_number(as.character(n_scales)))

n_scales_full <- dat_n_scales_num %>%
  summarise(
    n = sum(!is.na(n_scales_num)),
    mean = mean(n_scales_num, na.rm = TRUE),
    sd = sd(n_scales_num, na.rm = TRUE),
    median = median(n_scales_num, na.rm = TRUE),
    q25 = quantile(n_scales_num, 0.25, na.rm = TRUE),
    q75 = quantile(n_scales_num, 0.75, na.rm = TRUE),
    iqr = IQR(n_scales_num, na.rm = TRUE),
    min = min(n_scales_num, na.rm = TRUE),
    max = max(n_scales_num, na.rm = TRUE)
  )

n_scales_full

exclude_ids <- c("Newson2020", "Linton2016")

n_scales_excluded <- dat_n_scales_num %>%
  filter(!study_id %in% exclude_ids) %>%
  summarise(
    n = sum(!is.na(n_scales_num)),
    mean = mean(n_scales_num, na.rm = TRUE),
    sd = sd(n_scales_num, na.rm = TRUE),
    median = median(n_scales_num, na.rm = TRUE),
    q25 = quantile(n_scales_num, 0.25, na.rm = TRUE),
    q75 = quantile(n_scales_num, 0.75, na.rm = TRUE),
    iqr = IQR(n_scales_num, na.rm = TRUE),
    min = min(n_scales_num, na.rm = TRUE),
    max = max(n_scales_num, na.rm = TRUE)
  )


n_scales_excluded

```

```{r}
readr::write_csv(n_scales_full, here("outputs", "n_scales_full.csv"))

readr::write_csv(n_scales_excluded, here("outputs", "n_scales_excluded.csv"))
```


#Step 6: Conditional descriptives for overlap values (only where Jaccard values are reported)
```{r}
overlap_pairs <- tribble(
  ~flag, ~value,
  "mean_overlap", "mean_overlap_nr",
  "min_overlap",  "min_overlap_nr",
  "max_overlap",  "max_overlap_nr"
) %>%
  filter(flag %in% names(dat_num), value %in% names(dat_num))

overlap_conditional <- overlap_pairs %>%
  pmap_dfr(function(flag, value) {

    # 1) subset to rows where the metric is reported/applicable
    df_sub <- dat_num %>% filter(.data[[flag]] == TRUE)

    # 2) pull the numeric vector from the relevant *_nr column
    x <- df_sub %>% pull(.data[[value]])

    # 3) compute descriptives on the numeric vector
    summarise_numeric(x) %>%
      mutate(
        flag = flag,
        value = value,
        n_applicable = nrow(df_sub)  # how many rows satisfied flag == TRUE
      ) %>%
      relocate(flag, value, n_applicable)
  })

overlap_conditional

```

#Step 7: Percentage reduction from total items to overlap categories
From the 37 papers, only one author (Black et al., 2023) reported the percentage reduction from the total items of all scales to the symptom categories they ended up doing the content overlap with. So we calculated the percentage reduction for the remaining papers (wherever possible).
##Step 7a: Summary of overall percentage_reduction
```{r}
dat <- dat %>%
  mutate(percentage_reduction = as.numeric(percentage_reduction))

# 1) Overall summary of percentage_reduction
reduction_summary <- dat %>%
  summarise(
    n_total = n(),
    n_coded = sum(!is.na(percentage_reduction)),
    mean = mean(percentage_reduction, na.rm = TRUE),
    sd = sd(percentage_reduction, na.rm = TRUE),
    median = median(percentage_reduction, na.rm = TRUE),
    q25 = quantile(percentage_reduction, 0.25, na.rm = TRUE),
    q75 = quantile(percentage_reduction, 0.75, na.rm = TRUE),
    iqr = IQR(percentage_reduction, na.rm = TRUE),
    min = min(percentage_reduction, na.rm = TRUE),
    max = max(percentage_reduction, na.rm = TRUE)
  )

reduction_summary

```

##Step 7b: Comparison between approaches
There were two main approaches to categorizing the items; on the symptom_level vs. on the category/domain_level
```{r}

# - has_symptom: item_unit contains "symptom_level"
# - has_category: item_unit contains "category_level"

dat_units <- dat %>%
  mutate(
    item_unit_clean = item_unit %>%
      str_to_lower() %>%
      str_replace_all(",", ";") %>%      # guard against stray commas
      str_squish(),
    has_symptom  = str_detect(item_unit_clean, "\\bsymptom_level\\b"),
    has_category = str_detect(item_unit_clean, "\\bcategory_level\\b")
  )

symptom_stats <- dat_units %>%
  filter(has_symptom, !is.na(percentage_reduction)) %>%
  summarise(
    n = n(),
    mean = mean(percentage_reduction),
    median = median(percentage_reduction),
    q25 = quantile(percentage_reduction, 0.25),
    q75 = quantile(percentage_reduction, 0.75)
  ) %>%
  mutate(group = "symptom_level")

category_stats <- dat_units %>%
  filter(has_category, !is.na(percentage_reduction)) %>%
  summarise(
    n = n(),
    mean = mean(percentage_reduction),
    median = median(percentage_reduction),
    q25 = quantile(percentage_reduction, 0.25),
    q75 = quantile(percentage_reduction, 0.75)
  ) %>%
  mutate(group = "category_level")

unit_comparison <- bind_rows(symptom_stats, category_stats) %>%
  select(group, everything())

unit_comparison

```

## Step 7c: Descriptives for remaining item_unit levels 
```{r}
# Ensures consistent token formatting
dat_units2 <- dat %>%
  mutate(
    item_unit_clean = item_unit %>%
      str_to_lower() %>%
      str_replace_all(",", ";") %>%
      str_squish()
  )


# Descriptives for the remaining levels: semantic_level, criterion_level, item_level reductions 

other_tokens <- c("semantic_level", "criterion_level", "item_level")

reduction_other_levels <- dat_units2 %>%
  select(study_id, item_unit_clean, percentage_reduction) %>%
  separate_rows(item_unit_clean, sep = ";\\s*") %>%
  mutate(item_unit_clean = str_squish(item_unit_clean)) %>%
  filter(item_unit_clean %in% other_tokens) %>%
  group_by(item_unit_clean) %>%
  summarise(
    n_total = n(),
    n_coded = sum(!is.na(percentage_reduction)),
    mean = mean(percentage_reduction, na.rm = TRUE),
    sd = sd(percentage_reduction, na.rm = TRUE),
    median = median(percentage_reduction, na.rm = TRUE),
    q25 = quantile(percentage_reduction, 0.25, na.rm = TRUE),
    q75 = quantile(percentage_reduction, 0.75, na.rm = TRUE),
    min = min(percentage_reduction, na.rm = TRUE),
    max = max(percentage_reduction, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(item_unit_clean)

reduction_other_levels

```


ORGANIZED WORKFLOW FOR FACTORS + TAG COLUMNS
```{r}
# 1) Standardize selected string columns once
# 2) Use small helper functions to generate:
#    - frequency tables for factor-like columns (single value per row)
#    - token prevalence tables for tag columns (semicolon-separated values)
#    - token width -> how many tokens per study/row

# std_text(): 
# - Converts to character (in case Excel imported factors)
# - Lowercases
# - Trims and collapses whitespace
#
# std_tags():
# - Does std_text()
# - Normalizes delimiters so that:
#   - commas become semicolons
#   - "; " becomes ";"

std_text <- function(x) {
  x %>%
    as.character() %>%
    str_to_lower() %>%
    str_squish()
}

std_tags <- function(x) {
  x %>%
    std_text() %>%
    str_replace_all(",", ";") %>%     # treat commas as tag separators if they occur
    str_replace_all(";\\s*", ";")     # remove spaces after semicolons
}

# one vector to put all "single-value factor"- columns in
factor_cols_to_std <- c(
  "population",
  "overarching_domain",
  "administration_mode",
  "compound_items_codes",
  "categorization_approach",
  "software_used",
  "matrix_source"
)

# Selects all columns to be standardized as "tag columns"
tag_cols_to_std <- c(
  "database",
  "item_unit",
  "diagnostic_framework"
)

# Create a standardized working dataset without accidental commas or whitespaces

dat_std <- dat %>%
  mutate(
    across(any_of(factor_cols_to_std), std_text),
    across(any_of(tag_cols_to_std), std_tags)
  )

```

# Helper functions: frequency table for factor-like columns
```{r}
# 1) Function for frequency table for factor-like columns

# - Counts unique values of a column (one value per row)
# - Adds percent column relative to the total n counted

tab_factor <- function(data, col) {
  data %>%
    count({{ col }}, name = "count") %>%
    mutate(percent = scales::percent(count / sum(count), accuracy = 0.1)) %>%
    arrange(desc(count))
}

# 2) function for tokenizing  + counting tag prevalence

# for columns like:
# - database = "pubmed;psycinfo;google_scholar"

# - Splits tag strings into one row per (study_id, tag)
# - Removes "NA"/"indeterminable"/empty tokens
# - Counts in how many distinct studies each tag appears
# - Adds percent relative to total number of studies/rows

tab_tags_usage <- function(data, id_col, tags_col,
                           drop_vals = c("na", "")) {
  data %>%
    select({{ id_col }}, {{ tags_col }}) %>%
    separate_rows({{ tags_col }}, sep = ";") %>%
    mutate({{ tags_col }} := str_trim({{ tags_col }})) %>%
    filter(!is.na({{ tags_col }}), !({{ tags_col }} %in% drop_vals)) %>%
    count({{ tags_col }}, name = "n_papers") %>%
    mutate(percent = scales::percent(n_papers / n_distinct(data %>% pull({{ id_col }})),
                                     accuracy = 0.1)) %>%
    arrange(desc(n_papers))
}


# 3) function: how many tags per study

# - How many databases did papers use on average?
# - How many papers used only 1 database?
# - Distribution of number of tags per paper
#
tab_tags_breadth <- function(data, id_col, tags_col,
                             drop_vals = c("na", "", "indeterminable")) {
  data %>%
    select({{ id_col }}, {{ tags_col }}) %>%
    separate_rows({{ tags_col }}, sep = ";") %>%
    mutate({{ tags_col }} := str_trim({{ tags_col }})) %>%
    filter(!is.na({{ tags_col }}), !({{ tags_col }} %in% drop_vals)) %>%
    distinct({{ id_col }}, {{ tags_col }}) %>%
    count({{ id_col }}, name = "n_tags")
}

```


```{r}
# A) Factor-like columns (one value per row)

tab_population <- tab_factor(dat_std, population)
tab_overarching_domain <- tab_factor(dat_std, overarching_domain)
tab_administration_mode <- tab_factor(dat_std, administration_mode)
tab_compound_items_codes <- tab_factor(dat_std, compound_items_codes)
tab_categorization_approach <- tab_factor(dat_std, categorization_approach)
tab_software_used <- tab_factor(dat_std, software_used)
tab_matrix_source <- tab_factor(dat_std, matrix_source)


tab_population
tab_overarching_domain
tab_administration_mode
tab_compound_items_codes
tab_categorization_approach
tab_software_used
tab_matrix_source

# B) Tag columns: DATABASE

# 1) Which databases were used most often?
db_usage <- tab_tags_usage(dat_std, study_id, database)

# 2) How many databases per paper?
db_breadth <- tab_tags_breadth(dat_std, study_id, database)

# 3) Summaries for breadth
db_breadth_summary <- db_breadth %>%
  summarise(
    n_coded = n(),                 # how many studies had at least one database token
    mean = mean(n_tags),
    median = median(n_tags),
    min = min(n_tags),
    max = max(n_tags),
    n_single_db = sum(n_tags == 1) # how many used exactly 1 database
  )

db_usage
db_breadth_summary

# Distribution of number of databases (1,2,3,...)
db_breadth_dist <- db_breadth %>%
  count(n_tags, name = "n_papers") %>%
  mutate(percent = scales::percent(n_papers / sum(n_papers), accuracy = 0.1)) %>%
  arrange(n_tags)

db_breadth_dist



# C) Tag columns: ITEM UNIT

item_unit_usage <- tab_tags_usage(dat_std, study_id, item_unit)
item_unit_breadth <- tab_tags_breadth(dat_std, study_id, item_unit)

item_unit_usage

# how many papers have >1 unit token (proxy for "multiple analyses/units")
item_unit_multi <- item_unit_breadth %>%
  summarise(
    n_total = n(),
    n_multi = sum(n_tags > 1),
    percent_multi = scales::percent(n_multi / n_total, accuracy = 0.1)
  )

item_unit_multi


# D) Tag columns: DIAGNOSTIC FRAMEWORK

diag_fw_usage <- tab_tags_usage(dat_std, study_id, diagnostic_framework)
diag_fw_breadth <- tab_tags_breadth(dat_std, study_id, diagnostic_framework)

diag_fw_usage

diag_fw_breadth %>%
  summarise(
    n_coded = n(),
    mean = mean(n_tags),
    median = median(n_tags),
    min = min(n_tags),
    max = max(n_tags)
  )

```
```{r}
readr::write_csv(tab_population, here("outputs", "tab_population.csv"))

readr::write_csv(tab_overarching_domain, here("outputs", "tab_overarching_domain.csv"))

readr::write_csv(tab_administration_mode, here("outputs", "tab_administration_mode.csv"))

```

```{r}
install.packages(c("flextable", "officer"))
library(flextable)
library(officer)
library(dplyr)

```


```{r}

ft_minimal <- function(df, caption = NULL) {
  ft <- flextable(df)

  if (!is.null(caption)) ft <- set_caption(ft, caption)

  ft %>%
    theme_booktabs() %>%                 # minimalist academic look (horizontal rules)
    autofit() %>%                        # fit columns to content
    fontsize(size = 10, part = "all") %>%
    align(align = "left", part = "all") %>%
    valign(valign = "top", part = "all") %>%
    padding(padding = 2, part = "all") %>%
    set_table_properties(width = 1, layout = "autofit")
}

```

```{r}
db_usage_out <- db_usage %>%
  rename(Database = database, `Studies (n)` = n_papers, `Studies (%)` = percent)

db_breadth_dist_out <- db_breadth_dist %>%
  rename(`# Databases` = n_tags, `Studies (n)` = n_papers, `Studies (%)` = percent)

db_breadth_summary_out <- db_breadth_summary %>%
  mutate(across(where(is.numeric), ~round(.x, 2)))


```

```{r}
doc <- read_docx()

doc <- doc %>%
  body_add_par("Database reporting and search breadth", style = "heading 2") %>%
  body_add_flextable(ft_minimal(db_breadth_summary_out, "Table X. Summary of database breadth")) %>%
  body_add_par("", style = "Normal") %>%
  body_add_flextable(ft_minimal(db_breadth_dist_out, "Table Y. Number of databases reported per study")) %>%
  body_add_par("", style = "Normal") %>%
  body_add_flextable(ft_minimal(db_usage_out, "Table Z. Prevalence of reported databases/search sources"))

print(doc, target = here::here("outputs", "tables_database_section.docx"))

```

```{r}
# We work only among studies that did a scale search
dat_ss <- dat_binary %>% 
  filter(scale_search == TRUE)

# Denominator for conditional inclusion-criteria percentages:
# (how many scale-search studies reported inclusion criteria?)
n_incl <- dat_ss %>%
  summarise(n_incl = sum(inclusion_binary == TRUE, na.rm = TRUE)) %>%
  pull(n_incl)

# ------------------------------------------------------------
# A) TOP BLOCK: format of selection criteria (denominator = 32)
# ------------------------------------------------------------
criteria_combo_block <- dat_ss %>%
  # keep only the two "format" indicators
  transmute(inclusion_binary, exclusion_criteria) %>%
  # (if you ever have NA here and want to keep it, tell me—right now you said n_na=0)
  filter(!is.na(inclusion_binary), !is.na(exclusion_criteria)) %>%
  count(inclusion_binary, exclusion_criteria, name = "n_true") %>%
  mutate(
    n_coded = sum(n_true),
    variable = case_when(
      inclusion_binary == TRUE  & exclusion_criteria == TRUE  ~ "Selection criteria format: both inclusion & exclusion",
      inclusion_binary == TRUE  & exclusion_criteria == FALSE ~ "Selection criteria format: inclusion only",
      inclusion_binary == FALSE & exclusion_criteria == TRUE  ~ "Selection criteria format: exclusions only",
      inclusion_binary == FALSE & exclusion_criteria == FALSE ~ "Selection criteria format: neither inclusion nor exclusion",
      TRUE ~ "Selection criteria format: other"
    ),
    percent = scales::percent(n_true / n_coded, accuracy = 0.1),
    `Studies (%)` = paste0(n_true, "/", n_coded, " (", percent, ")")
  ) %>%
  select(variable, n_true, `Studies (%)`) %>%
  arrange(desc(n_true))

# ------------------------------------------------------------
# B) BOTTOM BLOCK: inclusion criteria content (denominator = n_incl = 27)
# Only consider studies that reported inclusion criteria (inclusion_binary == TRUE)
# ------------------------------------------------------------
incl_criteria_cols <- c(
  "clinical_relevance",
  "language_mentioned",
  "availability_mentioned",
  "release_date_mentioned",
  "frequency",
  "min_scale_citation"
)

criteria_inclusion_block <- dat_ss %>%
  filter(inclusion_binary == TRUE) %>%
  summarise(
    across(
      all_of(incl_criteria_cols),
      ~ sum(.x %in% TRUE, na.rm = TRUE),
      .names = "{.col}"
    )
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "n_true"
  ) %>%
  mutate(
    variable = recode(variable,
      language_mentioned     = "Inclusion criteria: language",
      frequency              = "Inclusion criteria: frequency",
      availability_mentioned = "Inclusion criteria: availability",
      release_date_mentioned = "Inclusion criteria: release date/recency",
      clinical_relevance     = "Inclusion criteria: clinical relevance",
      min_scale_citation     = "Inclusion criteria: minimum citation threshold"
    ),
    percent = scales::percent(n_true / n_incl, accuracy = 0.1),
    `Studies (%)` = paste0(n_true, "/", n_incl, " (", percent, ")")
  ) %>%
  select(variable, n_true, `Studies (%)`) %>%
  arrange(desc(n_true))



# --- Append as 4th table to the same Word document ---
doc <- doc %>%
  body_add_par("", style = "Normal") %>%
  body_add_flextable(
    ft_minimal(
      criteria_table_final,
      "Table W. Reported scale-selection criteria among studies conducting a scale search (N = 32)"
    )
  )

# --- Save/overwrite the same Word file (or choose a new filename) ---
dir.create(here("outputs"), showWarnings = FALSE)

print(doc, target = here("outputs", "tables_database_section.docx"))
```

