---
title: "Methods descriptives"
author: "Tim"
date: "2026-04-15"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

# Aim of this script

This script summarizes the methodological characteristics of the included item-content analyses (ICAs).
It follows the order of the results chapter and keeps the workflow as linear as possible: load the data, create a few clearly separated working datasets, then calculate the descriptive summaries used in the text, tables, and figures.

# Step 1: Load packages

The analysis is based mainly on `tidyverse`. `janitor` is used to clean the column names, `here` handles file paths, and `officer` together with `flextable` is used for a Word export of the main tables.

```{r packages}
library(tidyverse)
library(janitor)
library(here)
library(scales)
library(officer)
library(flextable)
```

# Step 2: Define a few small helper functions

Only a small number of helper functions are used. They are kept simple and are all based on vectors rather than on more abstract tidy-evaluation.

```{r helper-functions}
std_text <- function(x) {
  x <- as.character(x)
  x <- stringr::str_to_lower(x)
  x <- stringr::str_squish(x)
  x
}

std_tags <- function(x) {
  x <- std_text(x)
  x <- stringr::str_replace_all(x, ",", ";")
  x <- stringr::str_replace_all(x, ";\\s*", ";")
  x
}

to_logical_strict <- function(x, colname = "unknown column") {
  x_chr <- as.character(x)
  x_std <- std_text(x_chr)

  true_vals <- c("true", "t", "yes", "y", "1")
  false_vals <- c("false", "f", "no", "n", "0")

  out <- dplyr::case_when(
    is.na(x_std) ~ NA,
    x_std %in% true_vals ~ TRUE,
    x_std %in% false_vals ~ FALSE,
    TRUE ~ NA
  )

  bad_vals <- unique(x_chr[!is.na(x_std) & !(x_std %in% c(true_vals, false_vals))])

  if (length(bad_vals) > 0) {
    stop(
      paste0(
        "Unexpected non-boolean value(s) in ", colname, ": ",
        paste(bad_vals, collapse = ", ")
      )
    )
  }

  out
}

prop_text <- function(n, d, accuracy = 0.1) {
  if (is.na(d) || d == 0) return(NA_character_)
  paste0(n, "/", d, " (", scales::percent(n / d, accuracy = accuracy), ")")
}

numeric_summary <- function(x) {
  x <- x[!is.na(x)]

  if (length(x) == 0) {
    return(tibble(
      n = 0,
      mean = NA_real_,
      sd = NA_real_,
      median = NA_real_,
      iqr = NA_real_,
      min = NA_real_,
      max = NA_real_,
      q25 = NA_real_,
      q75 = NA_real_
    ))
  }

  tibble(
    n = length(x),
    mean = mean(x),
    sd = sd(x),
    median = median(x),
    iqr = IQR(x),
    min = min(x),
    max = max(x),
    q25 = unname(quantile(x, 0.25)),
    q75 = unname(quantile(x, 0.75))
  )
}

ft_minimal <- function(df, caption = NULL) {
  ft <- flextable(df)

  if (!is.null(caption)) {
    ft <- set_caption(ft, caption)
  }

  ft %>%
    theme_booktabs() %>%
    autofit() %>%
    fontsize(size = 10, part = "all") %>%
    align(align = "left", part = "all") %>%
    valign(valign = "top", part = "all")
}
```

# Step 3: Define paths and load raw data

The CSV is read in as character data first. This avoids accidental coercion of semicolon-delimited fields such as `item_unit` or `percentage_reduction`.

```{r load-data}
output_dir <- here::here("outputs")

if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}

path_in <- here::here("data", "ica_methods.csv")

raw <- readr::read_csv(
  file = path_in,
  col_types = readr::cols(.default = readr::col_character()),
  na = c("", "NA", "N/A", "na", "n/a", "null", "NULL")
)

dat_raw <- raw %>%
  janitor::clean_names()

glimpse(dat_raw)  # sanity check
```

```{r required-columns}
required_cols <- c(
  "study_id",
  "n_scales",
  "population",
  "overarching_domain",
  "administration_mode",
  "scale_search",
  "database",
  "prisma_a",
  "prisma",
  "inclusion_binary",
  "clinical_relevance",
  "language_mentioned",
  "availability_mentioned",
  "release_date_mentioned",
  "frequency",
  "min_scale_citation",
  "exclusion_criteria",
  "item_unit",
  "percentage_reduction",
  "categorization_approach",
  "diagnostic_framework",
  "within_scale_item_collapse",
  "reverse_worded_items",
  "item_exclusions",
  "compound_items",
  "compound_items_codes",
  "idiosyncratic_items",
  "coding_team_reported",
  "initial_coding_team_size",
  "n_rater",
  "rater_expertise_mentioned",
  "blinded",
  "irr_reported",
  "codebook",
  "software_mentioned",
  "preregistered_rules",
  "coding_scheme_independent_mentioned"
)

missing_required <- setdiff(required_cols, names(dat_raw))

if (length(missing_required) > 0) {
  stop(
    paste0(
      "The following required columns are missing: ",
      paste(missing_required, collapse = ", ")
    )
  )
}

if (nrow(dat_raw) != 37) {
  warning(
    paste0(
      "Expected 37 studies, but found ", nrow(dat_raw),
      ". Check whether the current CSV export has changed."
    )
  )
}
```

# Step 4: Create a few working datasets

To keep the workflow readable, the raw import is preserved and a few separate working datasets are created. The main idea is simple: one dataset for logical variables, one for numeric variables, and one for standardized text fields.

## Step 4a: Logical variables

```{r prepare-logical-data}
logical_cols <- c(
  "matrix",
  "short_scales",
  "scale_search",
  "prisma_a",
  "prisma",
  "inclusion_criteria_applicable",
  "inclusion_binary",
  "clinical_relevance",
  "language_mentioned",
  "availability_mentioned",
  "release_date_mentioned",
  "frequency",
  "min_scale_citation",
  "exclusion_criteria",
  "multiple_analyses",
  "within_scale_item_collapse",
  "compound_items",
  "idiosyncratic_items",
  "reverse_worded_items",
  "coding_team_reported",
  "rater_expertise_mentioned",
  "blinded",
  "codebook",
  "software_mentioned",
  "item_exclusions",
  "irr_reported",
  "preregistered_rules",
  "coding_scheme_independent_mentioned",
  "mean_overlap",
  "min_overlap",
  "max_overlap",
  "length_correlation_noted"
)

logical_cols_present <- intersect(logical_cols, names(dat_raw))

dat_logical <- dat_raw

for (col in logical_cols_present) {
  dat_logical[[col]] <- to_logical_strict(dat_logical[[col]], colname = col)
}
```

## Step 4b: Numeric variables

Only variables that are clearly numeric are parsed here. The semicolon-delimited reduction field is deliberately left untouched and handled later at the analysis level.

```{r prepare-numeric-data}
dat_num <- dat_raw %>%
  mutate(
    n_scales = readr::parse_number(n_scales),
    initial_coding_team_size = readr::parse_number(initial_coding_team_size),
    n_rater = readr::parse_number(n_rater),
    mean_overlap_nr = readr::parse_number(mean_overlap_nr),
    min_overlap_nr = readr::parse_number(min_overlap_nr),
    max_overlap_nr = readr::parse_number(max_overlap_nr)
  )
```

## Step 4c: Standardized text and tag fields

```{r prepare-text-data}
dat_text <- dat_raw %>%
  mutate(
    population = std_text(population),
    overarching_domain = std_text(overarching_domain),
    administration_mode = std_text(administration_mode),
    categorization_approach = std_text(categorization_approach),
    compound_items_codes = std_text(compound_items_codes),
    software_used = std_text(software_used),
    matrix_source = std_text(matrix_source),
    database = std_tags(database),
    item_unit = std_tags(item_unit),
    diagnostic_framework = std_tags(diagnostic_framework)
  )
```

# Step 5: Overview of included ICAs

This section summarizes the broad descriptive characteristics of the included studies.

## Step 5a: Domain, administration mode, and population

```{r overview-factor-tables}
tab_population <- dat_text %>%
  count(population, name = "n") %>%
  mutate(percent = scales::percent(n / sum(n), accuracy = 0.1)) %>%
  arrange(desc(n))

tab_overarching_domain <- dat_text %>%
  count(overarching_domain, name = "n") %>%
  mutate(percent = scales::percent(n / sum(n), accuracy = 0.1)) %>%
  arrange(desc(n))

tab_administration_mode <- dat_text %>%
  count(administration_mode, name = "n") %>%
  mutate(percent = scales::percent(n / sum(n), accuracy = 0.1)) %>%
  arrange(desc(n))

tab_population
tab_overarching_domain
tab_administration_mode
```

## Step 5b: Number of scales per study

A second summary excludes the two broad mapping studies with exceptionally large instrument sets.

```{r n-scales-descriptives}
n_scales_full <- numeric_summary(dat_num$n_scales)

exclude_ids <- c("Newson2020", "Linton2016")

n_scales_excluded <- dat_num %>%
  filter(!study_id %in% exclude_ids) %>%
  pull(n_scales) %>%
  numeric_summary()

n_scales_full
n_scales_excluded
```

# Step 6: Scale search and selection transparency

This section summarizes whether studies actively assembled instrument sets, whether PRISMA-type reporting was used when applicable, which information sources were reported, and how selection criteria were documented.

## Step 6a: Basic binary summaries relevant to the search process

```{r search-binary-summaries}
scale_search_n <- sum(dat_logical$scale_search == TRUE, na.rm = TRUE)
scale_search_d <- sum(!is.na(dat_logical$scale_search))

prisma_applicable_n <- sum(dat_logical$prisma_a == TRUE, na.rm = TRUE)
prisma_applicable_d <- sum(!is.na(dat_logical$prisma_a))

prisma_reported_n <- dat_logical %>%
  filter(prisma_a == TRUE) %>%
  summarise(n = sum(prisma == TRUE, na.rm = TRUE)) %>%
  pull(n)

prisma_reported_d <- dat_logical %>%
  filter(prisma_a == TRUE) %>%
  summarise(n = sum(!is.na(prisma))) %>%
  pull(n)

search_summary <- tibble(
  indicator = c("Scale search/manual selection", "PRISMA applicable", "PRISMA reported among applicable studies"),
  n = c(scale_search_n, prisma_applicable_n, prisma_reported_n),
  denominator = c(scale_search_d, prisma_applicable_d, prisma_reported_d),
  report = c(
    prop_text(scale_search_n, scale_search_d),
    prop_text(prisma_applicable_n, prisma_applicable_d),
    prop_text(prisma_reported_n, prisma_reported_d)
  )
)

search_summary
```

## Step 6b: Database reporting and search breadth

Databases are stored as semicolon-delimited fields. The code below first expands them into a long table and then summarizes both source prevalence and the number of sources per study.

```{r database-tables}
db_long <- dat_text %>%
  select(study_id, database) %>%
  separate_rows(database, sep = ";") %>%
  mutate(database = stringr::str_trim(database)) %>%
  filter(!is.na(database), database != "") %>%
  distinct(study_id, database)

db_usage <- db_long %>%
  count(database, name = "n") %>%
  mutate(percent = scales::percent(n / nrow(dat_text), accuracy = 0.1)) %>%
  arrange(desc(n), database)

db_breadth <- db_long %>%
  filter(database != "indeterminable") %>%
  count(study_id, name = "n_sources")

db_breadth_summary <- tibble(
  n_scale_search_studies = sum(dat_logical$scale_search == TRUE, na.rm = TRUE),
  n_with_database_info = nrow(db_breadth),
  mean = mean(db_breadth$n_sources),
  median = median(db_breadth$n_sources),
  min = min(db_breadth$n_sources),
  max = max(db_breadth$n_sources),
  n_single_database = sum(db_breadth$n_sources == 1)
)

db_breadth_distribution <- db_breadth %>%
  count(n_sources, name = "n_studies") %>%
  mutate(percent = scales::percent(n_studies / sum(n_studies), accuracy = 0.1)) %>%
  arrange(n_sources)

db_usage
db_breadth_summary
db_breadth_distribution
```

## Step 6c: Selection criteria format

This summary is conditional on studies that conducted a scale search.

```{r selection-criteria-format}
dat_search <- dat_logical %>%
  filter(scale_search == TRUE)

criteria_combo_block <- dat_search %>%
  filter(!is.na(inclusion_binary), !is.na(exclusion_criteria)) %>%
  mutate(
    selection_criteria_format = case_when(
      inclusion_binary == TRUE & exclusion_criteria == TRUE ~ "Both inclusion & exclusion",
      inclusion_binary == TRUE & exclusion_criteria == FALSE ~ "Inclusion only",
      inclusion_binary == FALSE & exclusion_criteria == TRUE ~ "Exclusions only",
      inclusion_binary == FALSE & exclusion_criteria == FALSE ~ "Neither inclusion nor exclusion"
    )
  ) %>%
  count(selection_criteria_format, name = "n") %>%
  mutate(
    `Proportion (n/N)` = paste0(n, "/", sum(n)),
    Percent = scales::percent(n / sum(n), accuracy = 0.1)
  ) %>%
  rename(`Selection criteria format` = selection_criteria_format) %>%
  arrange(desc(n))

criteria_combo_block
```

## Step 6d: Specific inclusion criteria

This summary is restricted further to studies that reported inclusion criteria.

```{r inclusion-criteria-table}
dat_inclusion <- dat_search %>%
  filter(inclusion_binary == TRUE)

n_inclusion_studies <- nrow(dat_inclusion)

criteria_inclusion_block <- tibble(
  criterion = c(
    "language",
    "frequency",
    "availability",
    "release date/recency",
    "clinical relevance",
    "minimum citation threshold"
  ),
  n = c(
    sum(dat_inclusion$language_mentioned == TRUE, na.rm = TRUE),
    sum(dat_inclusion$frequency == TRUE, na.rm = TRUE),
    sum(dat_inclusion$availability_mentioned == TRUE, na.rm = TRUE),
    sum(dat_inclusion$release_date_mentioned == TRUE, na.rm = TRUE),
    sum(dat_inclusion$clinical_relevance == TRUE, na.rm = TRUE),
    sum(dat_inclusion$min_scale_citation == TRUE, na.rm = TRUE)
  )
) %>%
  mutate(
    `Proportion (n/N)` = paste0(n, "/", n_inclusion_studies),
    Percent = scales::percent(n / n_inclusion_studies, accuracy = 0.1)
  ) %>%
  rename(`Inclusion criteria` = criterion) %>%
  arrange(desc(n))

criteria_inclusion_block
```

# Step 7: Unit of comparison and percentage reduction

This section is handled on the analysis level because some studies report more than one overlap analysis. The key point is that `item_unit` and `percentage_reduction` must be split into aligned tokens within each study.

## Step 7a: Expand to an analysis-level dataset

```{r analysis-level-data}
analysis_long <- dat_raw %>%
  transmute(
    study_id,
    item_unit = std_tags(item_unit),
    percentage_reduction
  ) %>%
  rowwise() %>%
  mutate(
    unit_list = list({
      x <- item_unit
      if (is.na(x) || x == "") character(0) else stringr::str_split(x, ";", simplify = FALSE)[[1]]
    }),
    reduction_list = list({
      x <- percentage_reduction
      if (is.na(x) || x == "") {
        numeric(0)
      } else {
        vals <- stringr::str_split(as.character(x), ";", simplify = FALSE)[[1]]
        vals <- stringr::str_trim(vals)
        vals <- stringr::str_replace_all(vals, ",", ".")
        readr::parse_number(vals)
      }
    }),
    n_units = length(unit_list),
    n_reduction = length(reduction_list)
  ) %>%
  ungroup()

length_mismatch <- analysis_long %>%
  filter(n_reduction > 0, n_units != n_reduction)

if (nrow(length_mismatch) > 0) {
  warning(
    paste0(
      "Some studies have unequal numbers of item_unit and percentage_reduction entries: ",
      paste(length_mismatch$study_id, collapse = ", ")
    )
  )
}

analysis_long <- analysis_long %>%
  select(study_id, unit_list, reduction_list) %>%
  mutate(row_id = row_number()) %>%
  unnest_longer(unit_list, indices_to = "idx") %>%
  rename(unit_token = unit_list) %>%
  left_join(
    analysis_long %>%
      select(study_id, reduction_list) %>%
      mutate(row_id = row_number()) %>%
      unnest_longer(reduction_list, indices_to = "idx") %>%
      rename(percentage_reduction_num = reduction_list),
    by = c("study_id", "row_id", "idx")
  ) %>%
  select(-row_id) %>%
  mutate(unit_token = stringr::str_trim(unit_token))

if (nrow(analysis_long) != 44) {
  warning(
    paste0(
      "Expected 44 analysis-level rows based on item_unit. Found ", nrow(analysis_long), "."
    )
  )
}

analysis_long
```

## Step 7b: Unit-of-comparison counts

```{r unit-counts}
unit_counts_total_analyses <- analysis_long %>%
  count(unit_token, name = "n") %>%
  mutate(percent = scales::percent(n / sum(n), accuracy = 0.1)) %>%
  arrange(desc(n))

unit_counts_total_analyses
```

## Step 7c: Percentage reduction summaries

```{r percentage-reduction-summaries}
reduction_summary <- analysis_long %>%
  filter(!is.na(percentage_reduction_num)) %>%
  pull(percentage_reduction_num) %>%
  numeric_summary() %>%
  mutate(n_total_analyses = nrow(analysis_long)) %>%
  relocate(n_total_analyses)

reduction_by_unit_token <- analysis_long %>%
  filter(!is.na(percentage_reduction_num)) %>%
  group_by(unit_token) %>%
  summarise(
    n = n(),
    mean = mean(percentage_reduction_num),
    sd = sd(percentage_reduction_num),
    median = median(percentage_reduction_num),
    iqr = IQR(percentage_reduction_num),
    min = min(percentage_reduction_num),
    max = max(percentage_reduction_num),
    q25 = unname(quantile(percentage_reduction_num, 0.25)),
    q75 = unname(quantile(percentage_reduction_num, 0.75)),
    .groups = "drop"
  ) %>%
  arrange(desc(n)) %>% 
  mutate(
    across(c(mean, sd, median, iqr), ~ round(.x, 2)),
    across(c(min, max, q25, q75), ~ round(.x, 1))
  )

reduction_summary
reduction_by_unit_token
```

## Step 7d: Plot the two most common unit types

```{r reduction-plot}
plot_reduction <- analysis_long %>%
  filter(unit_token %in% c("symptom_level", "category_level")) %>%
  filter(!is.na(percentage_reduction_num)) %>%
  mutate(unit_token = factor(unit_token, levels = c("symptom_level", "category_level")))

fig_reduction_by_unit <- ggplot(plot_reduction, aes(x = unit_token, y = percentage_reduction_num)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.15, height = 0, alpha = 0.7) +
  labs(x = NULL, y = "Percentage reduction") +
  theme_minimal(base_size = 12)

fig_reduction_by_unit

ggsave(
  filename = file.path(output_dir, "fig_percentage_reduction_by_unit.png"),
  plot = fig_reduction_by_unit,
  width = 6.5,
  height = 4.5,
  dpi = 300,
  bg = "white"
)
```

# Step 8: Categorization approach and diagnostic framework

The categorization approach is a single-value field. The diagnostic-framework field can contain more than one tag per study.

```{r categorization-framework}
tab_categorization_approach <- dat_text %>%
  count(categorization_approach, name = "n") %>%
  mutate(percent = scales::percent(n / sum(n), accuracy = 0.1)) %>%
  arrange(desc(n))

diag_long <- dat_text %>%
  select(study_id, diagnostic_framework) %>%
  separate_rows(diagnostic_framework, sep = ";") %>%
  mutate(diagnostic_framework = stringr::str_trim(diagnostic_framework)) %>%
  filter(!is.na(diagnostic_framework), diagnostic_framework != "") %>%
  distinct(study_id, diagnostic_framework)

diag_fw_usage <- diag_long %>%
  count(diagnostic_framework, name = "n") %>%
  mutate(percent = scales::percent(n / nrow(dat_text), accuracy = 0.1)) %>%
  arrange(desc(n), diagnostic_framework)

diag_fw_breadth <- diag_long %>%
  count(study_id, name = "n_frameworks")

diag_fw_breadth_summary <- tibble(
  n_coded = nrow(diag_fw_breadth),
  mean = mean(diag_fw_breadth$n_frameworks),
  median = median(diag_fw_breadth$n_frameworks),
  min = min(diag_fw_breadth$n_frameworks),
  max = max(diag_fw_breadth$n_frameworks)
)

tab_categorization_approach
diag_fw_usage
diag_fw_breadth_summary
```

# Step 9: Item preprocessing and transformation decisions

This section reports the prevalence of the main preprocessing and transformation decisions. The compound-item subrows use a conditional denominator restricted to studies that mentioned compound items.

```{r transformation-table}
n_total_studies <- nrow(dat_logical)
n_compound <- sum(dat_logical$compound_items == TRUE, na.rm = TRUE)

tab_transform_main <- tibble(
  Decision = c(
    "Within-scale item collapsing",
    "Reverse-worded items collapsed",
    "Item exclusions reported",
    "Idiosyncratic/unique items discussed",
    "Compound items discussed"
  ),
  n = c(
    sum(dat_logical$within_scale_item_collapse == TRUE, na.rm = TRUE),
    sum(dat_logical$reverse_worded_items == TRUE, na.rm = TRUE),
    sum(dat_logical$item_exclusions == TRUE, na.rm = TRUE),
    sum(dat_logical$idiosyncratic_items == TRUE, na.rm = TRUE),
    sum(dat_logical$compound_items == TRUE, na.rm = TRUE)
  )
) %>%
  mutate(
    `Proportion (n/N)` = paste0(n, "/", n_total_studies),
    Percent = scales::percent(n / n_total_studies, accuracy = 0.1)
  )

tab_compound_sub <- dat_text %>%
  select(study_id, compound_items_codes) %>%
  left_join(dat_logical %>% select(study_id, compound_items), by = "study_id") %>%
  filter(compound_items == TRUE) %>%
  filter(!is.na(compound_items_codes), compound_items_codes != "") %>%
  count(compound_items_codes, name = "n") %>%
  mutate(
    Decision = case_when(
      compound_items_codes == "multi_coded" ~ "  ↳ Compound handling: multi-coded",
      compound_items_codes == "mono_coded" ~ "  ↳ Compound handling: mono-coded",
      compound_items_codes == "other" ~ "  ↳ Compound handling: other",
      TRUE ~ "  ↳ Compound handling: other/unclear"
    ),
    `Proportion (n/N)` = paste0(n, "/", n_compound),
    Percent = scales::percent(n / n_compound, accuracy = 0.1)
  ) %>%
  select(Decision, n, `Proportion (n/N)`, Percent) %>%
  arrange(desc(n))

tab_transformations <- bind_rows(tab_transform_main, tab_compound_sub)

tab_transformations
```

# Step 10: Coding workflow, reliability, and reproducibility

This section summarizes the currently extracted coding-workflow variables and the additional reproducibility indicators already in the dataset.

## Step 10a: Coding team and rater descriptives

```{r coding-team-descriptives}
coding_team_summary <- tibble(
  Indicator = "Coding team reported",
  n = sum(dat_logical$coding_team_reported == TRUE, na.rm = TRUE),
  denominator = sum(!is.na(dat_logical$coding_team_reported))
) %>%
  mutate(report = prop_text(n, denominator))

rater_expertise_summary <- tibble(
  Indicator = "Rater expertise mentioned",
  n = sum(dat_logical$rater_expertise_mentioned == TRUE, na.rm = TRUE),
  denominator = sum(!is.na(dat_logical$rater_expertise_mentioned))
) %>%
  mutate(report = prop_text(n, denominator))

blinded_summary <- tibble(
  Indicator = "Blinded",
  n = sum(dat_logical$blinded == TRUE, na.rm = TRUE),
  denominator = sum(!is.na(dat_logical$blinded))
) %>%
  mutate(report = prop_text(n, denominator))

initial_team_summary <- numeric_summary(dat_num$initial_coding_team_size)
n_rater_summary <- numeric_summary(dat_num$n_rater)

coding_team_summary
rater_expertise_summary
blinded_summary
initial_team_summary
n_rater_summary

team_size_table <- tibble::tibble(
  metric = c("Initial coding team size", "Number of raters"),
  n = c(initial_team_summary$n, n_rater_summary$n),
  mean = c(initial_team_summary$mean, n_rater_summary$mean),
  sd = c(initial_team_summary$sd, n_rater_summary$sd),
  median = c(initial_team_summary$median, n_rater_summary$median),
  iqr = c(initial_team_summary$iqr, n_rater_summary$iqr),
  min = c(initial_team_summary$min, n_rater_summary$min),
  max = c(initial_team_summary$max, n_rater_summary$max)
) %>% 
 mutate(
    across(c(mean, sd, median, iqr), ~ round(.x, 2)),
    across(c(n, min, max), ~ round(.x, 0))
  )

team_size_table

initial_team_size_dist <- dat_num %>%
  filter(!is.na(initial_coding_team_size)) %>%
  count(initial_coding_team_size, name = "n") %>%
  mutate(
    percent = 100 * n / sum(n)
  ) %>%
  arrange(initial_coding_team_size) %>%
  mutate(
    percent = round(percent, 1)
  )

initial_team_size_dist

n_rater_size_dist <- dat_num %>%
  filter(!is.na(n_rater)) %>%
  count(n_rater, name = "n") %>%
  mutate(
    percent = 100 * n / sum(n)
  ) %>%
  arrange(n_rater) %>%
  mutate(
    percent = round(percent, 1)
  )

n_rater_size_dist

coding_team_expansion <- dat_num %>%
  filter(!is.na(initial_coding_team_size), !is.na(n_rater)) %>%
  mutate(additional_raters = n_rater - initial_coding_team_size)

team_expansion_dist <- coding_team_expansion %>%
  count(additional_raters, name = "n") %>%
  mutate(
    percent = 100 * n / sum(n)
  ) %>%
  arrange(additional_raters) %>%
  mutate(
    percent = round(percent, 1)
  )

team_expansion_binary <- coding_team_expansion %>%
  mutate(expanded = additional_raters > 0) %>%
  count(expanded, name = "n") %>%
  mutate(
    percent = 100 * n / sum(n)
  ) %>%
  mutate(
    percent = round(percent, 1)
  )

team_expansion_binary

# Combine the two distributions for plotting
team_size_plot <- bind_rows(
  initial_team_size_dist %>%
    transmute(
      team_size = initial_coding_team_size,
      n,
      percent,
      type = "Initial coding team"
    ),
  n_rater_size_dist %>%
    transmute(
      team_size = n_rater,
      n,
      percent,
      type = "Total raters involved"
    )
)

team_size_plot

team_expansion_summary <- dat_num %>%
  filter(!is.na(initial_coding_team_size), !is.na(n_rater)) %>%
  mutate(expanded = n_rater > initial_coding_team_size) %>%
  summarise(
    n_expanded = sum(expanded),
    n_total = n(),
    percent_expanded = round(100 * n_expanded / n_total, 1)
  )

team_expansion_summary

# Step 10x: Faceted bar plot of coding-team distributions -----------------

team_dist_plot <- bind_rows(
  initial_team_size_dist %>%
    transmute(
      team_size = initial_coding_team_size,
      n,
      type = "Initial coding team"
    ),
  n_rater_size_dist %>%
    transmute(
      team_size = n_rater,
      n,
      type = "Total raters involved"
    )
) %>%
  mutate(
    team_size = as.integer(team_size)
  )

all_team_sizes <- seq(
  from = min(team_dist_plot$team_size, na.rm = TRUE),
  to = max(team_dist_plot$team_size, na.rm = TRUE)
)

team_dist_plot <- team_dist_plot %>%
  tidyr::complete(
    type,
    team_size = all_team_sizes,
    fill = list(n = 0)
  )

p_team_dist <- ggplot(team_dist_plot, aes(x = factor(team_size), y = n)) +
  geom_col(width = 0.48) +
  geom_text(
    aes(label = ifelse(n == 0, "", n)),
    vjust = -0.3,
    size = 3.5
  ) +
  facet_wrap(~ type, ncol = 1) +
  scale_y_continuous(
    limits = c(0, 16),
    breaks = c(0, 5, 10, 15),
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    x = "Number of coders",
    y = "Number of studies"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "plain", size = 11),
    strip.background = element_blank(),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    plot.title = element_blank()
  )

p_team_dist

# Figure: comparison of coding team sizes
p_team_size <- ggplot(team_size_plot, aes(x = factor(team_size), y = n, fill = type)) +
  geom_col(position = "dodge") +
  labs(
    x = "Number of coders",
    y = "Number of studies",
    fill = NULL,
    title = "Coding team size across studies"
  ) +
  theme_minimal()

p_team_size



ggsave(
  filename = file.path(output_dir, "team_distribution_plot.png"),
  plot = p_team_dist,
  width = 6.5,
  height = 5.5,
  bg = "white"
)

team_transition_table <- dat_num %>%
  filter(!is.na(initial_coding_team_size), !is.na(n_rater)) %>%
  count(initial_coding_team_size, n_rater, name = "n") %>%
  arrange(initial_coding_team_size, n_rater)

team_transition_table
```

## Step 10b: Additional reproducibility indicators

```{r reliability-indicators}
reliability_table <- tibble(
  Indicator = c(
    "Independent coding scheme mentioned",
    "Interrater reliability reported",
    "Disagreement resolved mentioned"
  ),
  `n (Studies)` = c(
    sum(dat_logical$coding_scheme_independent_mentioned == TRUE, na.rm = TRUE),
    sum(dat_logical$irr_reported == TRUE, na.rm = TRUE),
    sum(dat_logical$disagreement_resolved_mentioned == TRUE, na.rm = TRUE)
  ),
  denominator = c(
    sum(!is.na(dat_logical$coding_scheme_independent_mentioned)),
    sum(!is.na(dat_logical$irr_reported)),
    sum(!is.na(dat_logical$disagreement_resolved_mentioned))
  )
) %>%
  mutate(
    `Proportion (n/N)` = paste0(`n (Studies)`, "/", denominator),
    Percent = 100 * `n (Studies)` / denominator
  ) %>%
  mutate(
    Percent = paste0(round(Percent, 1), "%")
  ) %>%
  select(Indicator, `n (Studies)`, `Proportion (n/N)`, Percent)

reliability_table

```
# Step 10c: Reproducibility and documentation

```{r documentation-indicators}
documentation_table <- tibble(
  Indicator = c(
    "Codebook reported",
    "Software mentioned",
    "Preregistered rules reported"
  ),
  `n (Studies)` = c(
    sum(dat_logical$codebook == TRUE, na.rm = TRUE),
    sum(dat_logical$software_mentioned == TRUE, na.rm = TRUE),
    sum(dat_logical$preregistered_rules == TRUE, na.rm = TRUE)
  ),
  denominator = c(
    sum(!is.na(dat_logical$codebook)),
    sum(!is.na(dat_logical$software_mentioned)),
    sum(!is.na(dat_logical$preregistered_rules))
  )
) %>%
  mutate(
    `Proportion (n/N)` = paste0(`n (Studies)`, "/", denominator),
    Percent = 100 * `n (Studies)` / denominator
  ) %>%
  mutate(
    Percent = paste0(round(Percent, 1), "%")
  ) %>%
  select(Indicator, `n (Studies)`, `Proportion (n/N)`, Percent)

documentation_table


# Software used

software_used_table <- dat_text %>%
  filter(!is.na(software_used), software_used != "not_mentioned") %>%
  count(software_used, name = "n (Studies)") %>%
  complete(
    software_used = c("r", "python", "no_code"),
    fill = list(`n (Studies)` = 0)
  ) %>%
  mutate(
    `Proportion (n/N)` = paste0(`n (Studies)`, "/", sum(`n (Studies)`)),
    Percent = 100 * `n (Studies)` / sum(`n (Studies)`)
  ) %>%
  mutate(
    Percent = paste0(round(Percent, 1), "%"),
    `Software used` = dplyr::recode(
      software_used,
      "r" = "R",
      "python" = "Python",
      "no_code" = "No code"
    )
  ) %>%
  arrange(desc(`n (Studies)`)) %>%
  select(`Software used`, `n (Studies)`, `Proportion (n/N)`, Percent)

software_used_table

```
```{r}
# Step 11a: Reporting of overlap outputs ----------------------------------

overlap_reporting_table <- tibble::tibble(
  Indicator = c(
    "Mean overlap reported",
    "Minimum overlap reported",
    "Maximum overlap reported",
    "Length × mean-overlap relation noted"
  ),
  `n (Studies)` = c(
    sum(dat_logical$mean_overlap == TRUE, na.rm = TRUE),
    sum(dat_logical$min_overlap == TRUE, na.rm = TRUE),
    sum(dat_logical$max_overlap == TRUE, na.rm = TRUE),
    sum(dat_logical$length_correlation_noted == TRUE, na.rm = TRUE)
  ),
  denominator = c(
    sum(!is.na(dat_logical$mean_overlap)),
    sum(!is.na(dat_logical$min_overlap)),
    sum(!is.na(dat_logical$max_overlap)),
    sum(!is.na(dat_logical$length_correlation_noted))
  )
) %>%
  mutate(
    `Proportion (n/N)` = paste0(`n (Studies)`, "/", denominator),
    Percent = 100 * `n (Studies)` / denominator
  ) %>%
  mutate(
    Percent = paste0(round(Percent, 1), "%")
  ) %>%
  select(Indicator, `n (Studies)`, `Proportion (n/N)`, Percent)

overlap_reporting_table

# Step 11b: Matrix source -------------------------------------------------

matrix_source_table <- dat_text %>%
  count(matrix_source, name = "n (Studies)") %>%
  mutate(
    `Proportion (n/N)` = paste0(`n (Studies)`, "/", sum(`n (Studies)`)),
    Percent = 100 * `n (Studies)` / sum(`n (Studies)`)
  ) %>%
  mutate(
    Percent = paste0(round(Percent, 1), "%"),
    `Matrix source` = dplyr::recode(
      matrix_source,
      "main_text" = "Main text",
      "supplement" = "Supplement",
      "not_available" = "Not available despite overlap reporting",
      .default = "Not applicable"
    )
  ) %>%
  arrange(match(`Matrix source`, c(
    "Main text",
    "Supplement",
    "Not available despite overlap reporting",
    "Not applicable"
  ))) %>%
  select(`Matrix source`, `n (Studies)`, `Proportion (n/N)`, Percent)

matrix_source_table
```

# Step 11c: Numeric descriptives of reported overlap summaries
```{r}

overlap_numeric_summary <- tibble::tibble(
  metric = c("Mean overlap", "Minimum overlap", "Maximum overlap"),
  n = c(
    sum(!is.na(dat_num$mean_overlap_nr)),
    sum(!is.na(dat_num$min_overlap_nr)),
    sum(!is.na(dat_num$max_overlap_nr))
  ),
  mean = c(
    mean(dat_num$mean_overlap_nr, na.rm = TRUE),
    mean(dat_num$min_overlap_nr, na.rm = TRUE),
    mean(dat_num$max_overlap_nr, na.rm = TRUE)
  ),
  median = c(
    median(dat_num$mean_overlap_nr, na.rm = TRUE),
    median(dat_num$min_overlap_nr, na.rm = TRUE),
    median(dat_num$max_overlap_nr, na.rm = TRUE)
  ),
  min = c(
    min(dat_num$mean_overlap_nr, na.rm = TRUE),
    min(dat_num$min_overlap_nr, na.rm = TRUE),
    min(dat_num$max_overlap_nr, na.rm = TRUE)
  ),
  max = c(
    max(dat_num$mean_overlap_nr, na.rm = TRUE),
    max(dat_num$min_overlap_nr, na.rm = TRUE),
    max(dat_num$max_overlap_nr, na.rm = TRUE)
  )
) %>%
  mutate(
    across(c(mean, median, min, max), ~ round(.x, 2))
  )

overlap_numeric_summary




# Step 11d: Types of length-overlap reporting -----------------------------

length_correlation_type_table <- dat_text %>%
  filter(!is.na(correlation_lengthxmeanoverlap_type)) %>%
  count(correlation_lengthxmeanoverlap_type, name = "n (Studies)") %>%
  mutate(
    `Proportion (n/N)` = paste0(`n (Studies)`, "/", sum(`n (Studies)`)),
    Percent = 100 * `n (Studies)` / sum(`n (Studies)`)
  ) %>%
  mutate(
    Percent = paste0(round(Percent, 1), "%")
  ) %>%
  arrange(desc(`n (Studies)`))

length_correlation_type_table
```

# Step 12a: Interpretation and recommendation indicators
```{r}

interpretation_indicators_table <- tibble(
  Indicator = c(
    "Heterogeneity emphasized",
    "Noninterchangeability emphasized",
    "Careful measure selection recommended",
    "Measure development recommended",
    "ICA methods problems/research needs noted"
  ),
  `n (Studies)` = c(
    sum(dat_logical$heterogeneity_emphasized == TRUE, na.rm = TRUE),
    sum(dat_logical$noninterchangeability_emphasized == TRUE, na.rm = TRUE),
    sum(dat_logical$recommend_careful_measure_selection == TRUE, na.rm = TRUE),
    sum(dat_logical$recommend_measure_development == TRUE, na.rm = TRUE),
    sum(dat_logical$ica_methods_problems == TRUE, na.rm = TRUE)
  ),
  denominator = c(
    sum(!is.na(dat_logical$heterogeneity_emphasized)),
    sum(!is.na(dat_logical$noninterchangeability_emphasized)),
    sum(!is.na(dat_logical$recommend_careful_measure_selection)),
    sum(!is.na(dat_logical$recommend_measure_development)),
    sum(!is.na(dat_logical$ica_methods_problems))
  )
) %>%
  mutate(
    `Proportion (n/N)` = paste0(`n (Studies)`, "/", denominator),
    Percent = 100 * `n (Studies)` / denominator
  ) %>%
  mutate(
    Percent = paste0(round(Percent, 1), "%")
  ) %>%
  select(Indicator, `n (Studies)`, `Proportion (n/N)`, Percent)

interpretation_indicators_table


# Step 12b: Reported overlap interpretation 

overlap_reported_table <- dat_text %>%
  filter(!is.na(overlap_reported)) %>%
  count(overlap_reported, name = "n (Studies)") %>%
  mutate(
    `Overlap reported` = dplyr::recode(
      overlap_reported,
      "very_weak_overlap" = "Very weak overlap",
      "weak_overlap" = "Weak overlap",
      "moderate_overlap" = "Moderate overlap",
      "not_specified" = "Not specified"
    )
  ) %>%
  arrange(match(`Overlap reported`, c(
    "Very weak overlap",
    "Weak overlap",
    "Moderate overlap",
    "Not specified"
  ))) %>%
  mutate(
    `Proportion (n/N)` = paste0(`n (Studies)`, "/", sum(`n (Studies)`)),
    Percent = 100 * `n (Studies)` / sum(`n (Studies)`)
  ) %>%
  mutate(
    Percent = paste0(round(Percent, 1), "%")
  ) %>%
  select(`Overlap reported`, `n (Studies)`, `Proportion (n/N)`, Percent)

overlap_reported_table
```


# Step 12: Export outputs

The main tables are written to the `outputs` folder as CSV files. A Word file is also created for convenience. The style selection for Word is kept defensive because style names can differ across local Word installations.

```{r write-csv-output}
readr::write_csv(tab_population, file.path(output_dir, "tab_population.csv"))
readr::write_csv(tab_overarching_domain, file.path(output_dir, "tab_overarching_domain.csv"))
readr::write_csv(tab_administration_mode, file.path(output_dir, "tab_administration_mode.csv"))
readr::write_csv(n_scales_full, file.path(output_dir, "n_scales_full.csv"))
readr::write_csv(n_scales_excluded, file.path(output_dir, "n_scales_excluded.csv"))
readr::write_csv(search_summary, file.path(output_dir, "search_summary.csv"))
readr::write_csv(db_usage, file.path(output_dir, "db_usage.csv"))
readr::write_csv(db_breadth_summary, file.path(output_dir, "db_breadth_summary.csv"))
readr::write_csv(db_breadth_distribution, file.path(output_dir, "db_breadth_distribution.csv"))
readr::write_csv(criteria_combo_block, file.path(output_dir, "criteria_combo_block.csv"))
readr::write_csv(criteria_inclusion_block, file.path(output_dir, "criteria_inclusion_block.csv"))
readr::write_csv(unit_counts_total_analyses, file.path(output_dir, "unit_counts_total_analyses.csv"))
readr::write_csv(reduction_summary, file.path(output_dir, "reduction_summary.csv"))
readr::write_csv(reduction_by_unit_token, file.path(output_dir, "reduction_by_unit_token.csv"))
readr::write_csv(tab_categorization_approach, file.path(output_dir, "tab_categorization_approach.csv"))
readr::write_csv(diag_fw_usage, file.path(output_dir, "diag_fw_usage.csv"))
readr::write_csv(tab_transformations, file.path(output_dir, "tab_transformations.csv"))
readr::write_csv(initial_team_size_dist, file.path(output_dir, "initial_team_size_dist.csv"))
readr::write_csv(team_expansion_binary, file.path(output_dir, "team_expansion_binary.csv"))
readr::write_csv(n_rater_size_dist, file.path(output_dir, "n_rater_size_dist.csv"))
readr::write_csv(reliability_table, file.path(output_dir, "reliability_indicators.csv"))







```

```{r write-word-output}
doc <- read_docx()

style_info <- styles_info(doc, type = "paragraph")
paragraph_styles <- style_info$style_name

heading1_style <- if ("Heading 1" %in% paragraph_styles) {
  "Heading 1"
} else if ("heading 1" %in% paragraph_styles) {
  "heading 1"
} else {
  "Normal"
}

heading2_style <- if ("Heading 2" %in% paragraph_styles) {
  "Heading 2"
} else if ("heading 2" %in% paragraph_styles) {
  "heading 2"
} else {
  "Normal"
}

normal_style <- if ("Normal" %in% paragraph_styles) {
  "Normal"
} else {
  paragraph_styles[1]
}

doc <- doc %>%
  body_add_par("Methods synthesis tables", style = heading1_style) %>%
  body_add_par("Overview", style = heading2_style) %>%
  body_add_flextable(ft_minimal(tab_population, "Population")) %>%
  body_add_par("", style = normal_style) %>%
  body_add_flextable(ft_minimal(tab_overarching_domain, "Overarching domain")) %>%
  body_add_par("", style = normal_style) %>%
  body_add_flextable(ft_minimal(tab_administration_mode, "Administration mode")) %>%
  body_add_par("", style = normal_style) %>%
  body_add_par("Scale search and selection transparency", style = heading2_style) %>%
  body_add_flextable(ft_minimal(db_usage, "Reported databases/search sources")) %>% 
  body_add_par("", style = normal_style) %>%
  body_add_flextable(ft_minimal(criteria_combo_block, "Selection criteria format")) %>%
  body_add_par("", style = normal_style) %>%
  body_add_flextable(ft_minimal(criteria_inclusion_block, "Specific inclusion criteria")) %>%
  body_add_par("", style = normal_style) %>%
  body_add_par("Unit of comparison and transformations", style = heading2_style) %>%
  body_add_flextable(ft_minimal(unit_counts_total_analyses, "Unit of comparison across analyses")) %>%
  body_add_par("", style = normal_style) %>%
  body_add_flextable(ft_minimal(reduction_by_unit_token, "Percentage reduction by unit type")) %>%
  body_add_par("", style = normal_style) %>%
  body_add_flextable(ft_minimal(tab_categorization_approach, "Categorization approach")) %>%
  body_add_par("", style = normal_style) %>%
  body_add_flextable(ft_minimal(diag_fw_usage, "Diagnostic frameworks referenced")) %>%
  body_add_par("", style = normal_style) %>%
  body_add_flextable(ft_minimal(tab_transformations, "Item preprocessing and transformation decisions")) %>%
  body_add_par("", style = normal_style) %>%
  body_add_par("Coding workflow and reproducibility", style = heading2_style) %>%
  body_add_par("", style = normal_style) %>%
  body_add_flextable(ft_minimal(team_size_table, "Coding team and rater counts")) %>% 
  body_add_par("", style = normal_style) %>%
  body_add_flextable(ft_minimal(initial_team_size_dist, "Initial team distribution")) %>% 
  body_add_par("", style = normal_style) %>%
  body_add_flextable(ft_minimal(n_rater_size_dist, "Rater team distribution")) %>% 
  body_add_par("", style = normal_style) %>%
  body_add_img(
  src = file.path(output_dir, "team_distribution_plot.png"),
  width = 6.2,
  height = 5.2
  ) %>%
  body_add_par("", style = normal_style) %>%
  body_add_flextable(ft_minimal(team_expansion_binary, "Coding team expansion")) %>%
  body_add_par("", style = normal_style) %>%
  body_add_flextable(ft_minimal(reliability_table, "Coding reliability indicators")) %>% 
  body_add_flextable(ft_minimal(documentation_table, "Documentation indicators")) %>%
  body_add_par("", style = normal_style) %>%
  body_add_flextable(ft_minimal(software_used_table, "Software used")) %>%
  body_add_par("", style = normal_style) %>% 
  body_add_flextable(ft_minimal(overlap_reporting_table, "Overlap reporting and scale length correlation")) %>%
  body_add_par("", style = normal_style) %>% 
  body_add_flextable(ft_minimal(matrix_source_table, "Location of overlap matrix")) %>%
  body_add_par("", style = normal_style) %>% 
  body_add_flextable(ft_minimal(length_correlation_type_table, "Types of reported length correlations")) %>%
  body_add_par("", style = normal_style) %>% 
  body_add_flextable(ft_minimal(interpretation_indicators_table, "Interpretive claims and recommendations")) %>%
  body_add_par("", style = normal_style) %>%
  body_add_flextable(ft_minimal(overlap_reported_table, "How authors described overlap strength")) %>%
  body_add_par("", style = normal_style)
  
  

print(doc, target = file.path(output_dir, "tables_methods_synthesis_simpler.docx"))
```

# Step 12: Notes for later extension

The remaining result sections, the discussion, and the limitations can be added later as additional steps. Because the raw data, logical variables, numeric variables, and analysis-level data are kept separate, new sections can be added without having to rewrite the earlier parts.
