---
title: "02_methods_descriptives"
author: "Tim"
date: "2026-01-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(janitor)
library(rlang)
library(scales)# for percent formatting
library(here) # robust paths within the repo
library(stringr)
```

```{r}
path_in <- here("data", "ica_methods.csv")

dir.create(here("outputs"), showWarnings = FALSE)

raw <- readr::read_csv(
file = path_in,
col_types = readr::cols(.default = readr::col_character()),
na = c("", "NA", "N/A", "na", "n/a", "null", "NULL")
)

dat <- raw %>% clean_names()

glimpse(dat) #sanity check
```

```{r}

binary_cols <- c(
  "matrix",
  "short_scales",
  "scale_search",
  "prisma_a",
  "prisma",
  "inclusion_criteria_applicable",
  "inclusion_binary",
  "clinical_relevance",
  "language_mentioned",
  "availability_mentioned",
  "release_date_mentioned",
  "frequency",
  "min_scale_citation",
  "exclusion_criteria",
  "multiple_analyses",
  "within_scale_item_collapse",
  "compound_items",
  "idiosyncratic_items",
  "reverse_worded_items",
  "coding_team_reported",
  "blinded",
  "codebook",
  "software_mentioned",
  "item_exclusions",
  "irr_reported",
  "preregistered_rules",
  "coding_scheme_independent_mentioned",
  "mean_overlap",
  "min_overlap",
  "max_overlap",
  "length_correlation_noted"
)

# Select only binary columns
dat_binary_raw <- dat %>%
  select(study_id, any_of(binary_cols))
``` 

```{r}
to_logical_strict <- function(x, colname) {
  # Leave logical as-is
  if (is.logical(x)) return(x)

  x0 <- as.character(x)
  x1 <- tolower(trimws(x0))

  true_vals  <- c("true", "t", "yes", "y", "1")
  false_vals <- c("false", "f", "no", "n", "0")

  out <- dplyr::case_when(
    is.na(x1) ~ NA,
    x1 %in% true_vals  ~ TRUE,
    x1 %in% false_vals ~ FALSE,
    TRUE ~ NA
  )

  # In case of typos: any non-NA value not in the allowed encodings
  bad <- which(!is.na(x1) & !(x1 %in% c(true_vals, false_vals)))
  if (length(bad) > 0) {
    bad_vals <- sort(unique(x0[bad]))
    rlang::abort(paste0(
      "Non-boolean value(s) detected in column '", colname, "': ",
      paste(bad_vals, collapse = ", "),
      ". Example row indices: ", paste(head(bad, 10), collapse = ", ")
    ))
  }

  out
}

dat_binary <- dat_binary_raw %>%
  mutate(across(all_of(binary_cols), ~ to_logical_strict(.x, cur_column())))

```

```{r}
binary_counts <- dat_binary %>%
  pivot_longer(  #long table with one row per (paper × variable)
    cols = all_of(binary_cols),
    names_to = "variable",
    values_to = "value"
  ) %>%
  mutate(value = case_when( #converts logicals into `"TRUE"/"FALSE"/"NA"` labels
    is.na(value) ~ "NA",
    value ~ "TRUE",
    !value ~ "FALSE"
  )) %>%
  count(variable, value, name = "n") %>%  #counts how many of each label occurs per variable
  group_by(variable) %>%
  mutate(n_total = sum(n)) %>% #computes the total number of observations per variable while still in long format
  ungroup() %>%
  pivot_wider(  #spreads the counts into columns 'TRUE', 'FALSE' and 'NA' and keeps the n_total
    names_from = value,
    values_from = n,
    values_fill = 0
  ) %>%
  arrange(desc(TRUE))

binary_counts
```

```{r}
binary_overall <- dat_binary %>%
  summarise(  #computes per-column counts
    across(
      all_of(binary_cols),
      list(
        n_true  = ~ sum(.x %in% TRUE, na.rm = TRUE),
        n_false = ~ sum(.x %in% FALSE, na.rm = TRUE),
        n_na    = ~ sum(is.na(.x)),
        n_coded = ~ sum(!is.na(.x))
      ),
      .names = "{.col}__{.fn}"  #attempt to give systematic names like "irr_reported_n_true"
    )
  ) %>%
  pivot_longer(  #splits the names into 'variable" (the original column name) and '.value' (which statistic the number belongs to)
    cols = everything(),
    names_to = c("variable", ".value"),
    names_sep = "__"
  ) %>%
  mutate(
    prop_true = n_true / n_coded,  #proportion TRUE among all coded cases only
    percent_true = scales::percent(prop_true, accuracy = 0.1),
    report = paste0(n_true, "/", n_coded, " (", percent_true, ")")  #manuscript-friendly string
  ) %>%
  arrange(desc(prop_true))

binary_overall %>% select(variable, n_true, n_coded, n_na, report)

readr::write_csv(binary_overall, here("outputs", "binary_overall.csv"))

```

# Step 4: Conditional reporting rates for gated variables
There are a couple of key "gates" in the binary dataset: 
1. `prisma` only matters when `prisma_a == TRUE`
2. “selection criteria” only matter when `scale_search == TRUE`

This is done as: **filter to applicable subset → summarise the target variables**

## Step 4a: PRISMA reporting among applicable studies (`prisma_a == TRUE`)
### Why this step?
It enables to report a conditional statement and avoids penalising studies where PRISMA is structurally not applicable:

> “Among studies that conducted their own scale search (PRISMA applicable), X% reported a PRISMA-style flow diagram.”

```{r}
prisma_cond <- dat_binary %>%
  filter(prisma_a == TRUE) %>%
  summarise(
    n_applicable = n(),
    n_true  = sum(prisma %in% TRUE, na.rm = TRUE),
    n_false = sum(prisma %in% FALSE, na.rm = TRUE),
    n_na    = sum(is.na(prisma)),
    n_coded = sum(!is.na(prisma))
  ) %>%
  mutate(
    prop_true = n_true / n_coded,
    percent_true = scales::percent(prop_true, accuracy = 0.1),
    report = paste0(n_true, "/", n_coded, " (", percent_true, ")")
  )

prisma_cond

```


## Step 4b: Selection-criteria reporting among manual searches (`scale_search == TRUE`)

```{r}
criteria_cols <- c(
  "inclusion_binary",
  "clinical_relevance",
  "language_mentioned",
  "availability_mentioned",
  "release_date_mentioned",
  "frequency",
  "min_scale_citation",
  "exclusion_criteria"
)

criteria_cond <- dat_binary %>%
  filter(scale_search == TRUE) %>%
  summarise(
    across(
      all_of(criteria_cols),
      list(
        n_true  = ~ sum(.x %in% TRUE, na.rm = TRUE),
        n_false = ~ sum(.x %in% FALSE, na.rm = TRUE),
        n_na    = ~ sum(is.na(.x)),
        n_coded = ~ sum(!is.na(.x))
      ),
      .names = "{.col}__{.fn}"
    )
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = c("variable", ".value"),
    names_sep = "__"
  ) %>%
  mutate(
    prop_true = n_true / n_coded,
    percent_true = scales::percent(prop_true, accuracy = 0.1),
    report = paste0(n_true, "/", n_coded, " (", percent_true, ")"),
    condition = "Among studies with manual scale search (scale_search == TRUE)"
  ) %>%
  arrange(desc(prop_true))

criteria_cond %>% select(condition, variable, n_true, n_coded, n_na, report)

```

```{r}
readr::write_csv(prisma_cond, here("outputs", "prisma_conditional.csv"))
readr::write_csv(criteria_cond, here("outputs", "criteria_conditional.csv"))

```

#Step 5: Descriptives for "homogenous" numeric columns

```{r}
# "clean"/homogenous numeric columns 
numeric_cols_clean <- c(
  "n_scales",
  "initial_coding_team_size",
  "n_rater",
  "mean_overlap_nr",
  "min_overlap_nr",
  "max_overlap_nr"
)

numeric_cols_clean_present <- intersect(numeric_cols_clean, names(dat))

dat_num <- dat %>%
  mutate(across(all_of(numeric_cols_clean_present), readr::parse_number))

# NA rates after parsing (quick QA)
numeric_na_rates <- dat_num %>%
  summarise(across(all_of(numeric_cols_clean_present),
                   ~ mean(is.na(.x)))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "prop_na") %>%
  mutate(percent_na = scales::percent(prop_na, accuracy = 0.1)) %>%
  arrange(desc(prop_na))

numeric_na_rates

``` 

```{r}
summarise_numeric <- function(x) {
  x <- x[!is.na(x)]
  if (length(x) == 0) {
    return(tibble(
      n = 0, mean = NA_real_, sd = NA_real_, median = NA_real_, iqr = NA_real_,
      min = NA_real_, max = NA_real_, q25 = NA_real_, q75 = NA_real_,
      n_outliers_iqr = NA_integer_
    ))
  }

  q <- quantile(x, probs = c(.25, .75), names = FALSE)
  iqr_val <- q[2] - q[1]
  lower <- q[1] - 1.5 * iqr_val
  upper <- q[2] + 1.5 * iqr_val
  n_out <- sum(x < lower | x > upper)

  tibble(
    n = length(x),
    mean = mean(x),
    sd = sd(x),
    median = median(x),
    iqr = IQR(x),
    min = min(x),
    max = max(x),
    q25 = q[1],
    q75 = q[2],
    n_outliers_iqr = n_out
  )
}

numeric_overall <- dat_num %>%
  summarise(across(all_of(numeric_cols_clean_present), summarise_numeric)) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "stats") %>%
  unnest_wider(stats)

numeric_overall

```

```{r}
readr::write_csv(numeric_overall, here("outputs", "numeric_overall.csv"))

```


#Step 6: Conditional descriptives for overlap values (only where Jaccard values are reported)
```{r}
overlap_pairs <- tribble(
  ~flag, ~value,
  "mean_overlap", "mean_overlap_nr",
  "min_overlap",  "min_overlap_nr",
  "max_overlap",  "max_overlap_nr"
) %>%
  filter(flag %in% names(dat_num), value %in% names(dat_num))

overlap_conditional <- overlap_pairs %>%
  pmap_dfr(function(flag, value) {

    # 1) subset to rows where the metric is reported/applicable
    df_sub <- dat_num %>% filter(.data[[flag]] == TRUE)

    # 2) pull the numeric vector from the relevant *_nr column
    x <- df_sub %>% pull(.data[[value]])

    # 3) compute descriptives on the numeric vector
    summarise_numeric(x) %>%
      mutate(
        flag = flag,
        value = value,
        n_applicable = nrow(df_sub)  # how many rows satisfied flag == TRUE
      ) %>%
      relocate(flag, value, n_applicable)
  })

overlap_conditional

```

#Step 7: Percentage reduction from total items to overlap categories
From the 37 papers, only one author (Black et al., 2023) reported the percentage reduction from the total items of all scales to the symptom categories they ended up doing the content overlap with. So we calculated the percentage reduction for the remaining papers (wherever possible).
##Step 7a: Summary of overall percentage_reduction
```{r}
dat <- dat %>%
  mutate(percentage_reduction = as.numeric(percentage_reduction))

# 1) Overall summary of percentage_reduction
reduction_summary <- dat %>%
  summarise(
    n_total = n(),
    n_coded = sum(!is.na(percentage_reduction)),
    mean = mean(percentage_reduction, na.rm = TRUE),
    sd = sd(percentage_reduction, na.rm = TRUE),
    median = median(percentage_reduction, na.rm = TRUE),
    q25 = quantile(percentage_reduction, 0.25, na.rm = TRUE),
    q75 = quantile(percentage_reduction, 0.75, na.rm = TRUE),
    iqr = IQR(percentage_reduction, na.rm = TRUE),
    min = min(percentage_reduction, na.rm = TRUE),
    max = max(percentage_reduction, na.rm = TRUE)
  )

reduction_summary

```

##Step 7b: Comparison between approaches
There were two main approaches to categorizing the items; on the symptom_level vs. on the category/domain_level
```{r}

# - has_symptom: item_unit contains "symptom_level"
# - has_category: item_unit contains "category_level"

dat_units <- dat %>%
  mutate(
    item_unit_clean = item_unit %>%
      str_to_lower() %>%
      str_replace_all(",", ";") %>%      # guard against stray commas
      str_squish(),
    has_symptom  = str_detect(item_unit_clean, "\\bsymptom_level\\b"),
    has_category = str_detect(item_unit_clean, "\\bcategory_level\\b")
  )

symptom_stats <- dat_units %>%
  filter(has_symptom, !is.na(percentage_reduction)) %>%
  summarise(
    n = n(),
    mean = mean(percentage_reduction),
    median = median(percentage_reduction),
    q25 = quantile(percentage_reduction, 0.25),
    q75 = quantile(percentage_reduction, 0.75)
  ) %>%
  mutate(group = "symptom_level")

category_stats <- dat_units %>%
  filter(has_category, !is.na(percentage_reduction)) %>%
  summarise(
    n = n(),
    mean = mean(percentage_reduction),
    median = median(percentage_reduction),
    q25 = quantile(percentage_reduction, 0.25),
    q75 = quantile(percentage_reduction, 0.75)
  ) %>%
  mutate(group = "category_level")

unit_comparison <- bind_rows(symptom_stats, category_stats) %>%
  select(group, everything())

unit_comparison

```

## Step 7c: Descriptives for remaining item_unit levels 
```{r}
# Ensures consistent token formatting
dat_units2 <- dat %>%
  mutate(
    item_unit_clean = item_unit %>%
      str_to_lower() %>%
      str_replace_all(",", ";") %>%
      str_squish()
  )


# Descriptives for the remaining levels: semantic_level, criterion_level, item_level reductions 

other_tokens <- c("semantic_level", "criterion_level", "item_level")

reduction_other_levels <- dat_units2 %>%
  select(study_id, item_unit_clean, percentage_reduction) %>%
  separate_rows(item_unit_clean, sep = ";\\s*") %>%
  mutate(item_unit_clean = str_squish(item_unit_clean)) %>%
  filter(item_unit_clean %in% other_tokens) %>%
  group_by(item_unit_clean) %>%
  summarise(
    n_total = n(),
    n_coded = sum(!is.na(percentage_reduction)),
    mean = mean(percentage_reduction, na.rm = TRUE),
    sd = sd(percentage_reduction, na.rm = TRUE),
    median = median(percentage_reduction, na.rm = TRUE),
    q25 = quantile(percentage_reduction, 0.25, na.rm = TRUE),
    q75 = quantile(percentage_reduction, 0.75, na.rm = TRUE),
    min = min(percentage_reduction, na.rm = TRUE),
    max = max(percentage_reduction, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(item_unit_clean)

reduction_other_levels

```

```{r}
tab_overarching_domain <- dat %>%
  select(overarching_domain) %>%
  count(overarching_domain, name = "n") %>%
  mutate(percent = scales::percent(n / sum(n), 0.1)) %>%
  arrange(desc(n))

tab_overarching_domain

```

```{r}
tab_compound_items_codes <- dat %>%
  select(compound_items_codes) %>%
  count(compound_items_codes, name = "n") %>%
  mutate(percent = scales::percent(n / sum(n), accuracy = 0.1)) %>%
  arrange(desc(n))

tab_compound_items_codes

```

```{r}
tab_categorization_approach <- dat %>%
  select(categorization_approach) %>%
  count(categorization_approach, name = "n") %>%
  mutate(percent = scales::percent(n / sum(n), accuracy = 0.1)) %>%
  arrange(desc(n))

tab_categorization_approach

```

```{r}
# ============================================================
# ORGANIZED WORKFLOW FOR FACTORS + TAG COLUMNS (TIDYVERSE)
# ============================================================
# Goal:
# 1) Standardize selected string columns once (optional but recommended).
# 2) Use small helper functions to generate:
#    - frequency tables for factor-like columns (single value per row)
#    - token prevalence tables for tag columns (semicolon-separated values)
#    - token breadth summaries (how many tokens per study/row)
#
# Assumptions:
# - Your dataset is called `dat`
# - `study_id` uniquely identifies each row (or at least each unit you want to count)
# - Tag columns use ";" as delimiter (e.g., "pubmed; psycinfo")
# - You want tables for supplementary materials (simple counts + percents)
# ============================================================



# ------------------------------------------------------------
# 0) OPTIONAL: Standardize only the columns you will analyze
# ------------------------------------------------------------
# Why do this?
# - It prevents accidental category-splitting due to formatting differences
#   (e.g., "PubMed" vs "pubmed", or "web of science" vs "web_of_science").
# If you are 100% sure your sheet is already standardized, you can skip this
# and just set dat_std <- dat.
#
# std_text(): 
# - Converts to character (in case Excel imported factors)
# - Lowercases
# - Trims and collapses whitespace
#
# std_tags():
# - Does std_text()
# - Normalizes delimiters so that:
#   - commas become semicolons
#   - "; " becomes ";"
#
std_text <- function(x) {
  x %>%
    as.character() %>%
    str_to_lower() %>%
    str_squish()
}

std_tags <- function(x) {
  x %>%
    std_text() %>%
    str_replace_all(",", ";") %>%     # treat commas as tag separators if they occur
    str_replace_all(";\\s*", ";")     # remove spaces after semicolons
}

# Choose which columns should be standardized as "single-value factors"
factor_cols_to_std <- c(
  "overarching_domain",
  "administration_mode",
  "compound_items_codes",
  "categorization_approach"
)

# Choose which columns should be standardized as "tag columns"
tag_cols_to_std <- c(
  "database",
  "item_unit",
  "diagnostic_framework"
)

# Create a standardized working dataset:
# - only the specified columns are transformed
# - all other columns are kept unchanged
dat_std <- dat %>%
  mutate(
    across(any_of(factor_cols_to_std), std_text),
    across(any_of(tag_cols_to_std), std_tags)
  )

# ------------------------------------------------------------
# 1) Helper function: frequency table for factor-like columns
# ------------------------------------------------------------
# What it does:
# - Counts unique values of a column (one value per row)
# - Adds percent column relative to the total n counted
#
# Why it is useful:
# - You can reuse it across many columns with a single line per column.
#
tab_factor <- function(data, col) {
  data %>%
    # count() creates a frequency table; `name` sets the count column name
    count({{ col }}, name = "count") %>%
    # percent is count divided by total count across categories
    mutate(percent = scales::percent(count / sum(count), accuracy = 0.1)) %>%
    # sort descending so most frequent category is on top
    arrange(desc(count))
}

# ------------------------------------------------------------
# 2) Helper function: tokenize + count tag prevalence
# ------------------------------------------------------------
# This is for columns like:
# - database = "pubmed;psycinfo;google_scholar"
#
# What it does:
# - Splits tag strings into one row per (study_id, tag)
# - Removes "NA"/"indeterminable"/empty tokens (configurable)
# - Counts in how many distinct studies each tag appears
# - Adds percent relative to total number of studies/rows
#
tab_tags_usage <- function(data, id_col, tags_col,
                           drop_vals = c("na", "indeterminable", "")) {
  data %>%
    # keep only the ID + the tag column to keep the pipeline clear
    select({{ id_col }}, {{ tags_col }}) %>%
    # separate_rows turns "a;b;c" into three rows (a), (b), (c) for the same ID
    separate_rows({{ tags_col }}, sep = ";") %>%
    # tidy tag tokens (e.g., remove accidental spaces)
    mutate({{ tags_col }} := str_trim({{ tags_col }})) %>%
    # drop missing or placeholder tokens
    filter(!is.na({{ tags_col }}), !({{ tags_col }} %in% drop_vals)) %>%
    # ensure each (ID, tag) pair is counted once even if duplicated
    distinct({{ id_col }}, {{ tags_col }}) %>%
    # count how many IDs used each tag
    count({{ tags_col }}, name = "n_papers") %>%
    # percent relative to the total number of distinct IDs in the dataset
    mutate(percent = scales::percent(n_papers / n_distinct(data %>% pull({{ id_col }})),
                                     accuracy = 0.1)) %>%
    arrange(desc(n_papers))
}

# ------------------------------------------------------------
# 3) Helper function: tag breadth (how many tags per study)
# ------------------------------------------------------------
# Useful questions it answers:
# - How many databases did papers use on average?
# - How many papers used only 1 database?
# - Distribution of number of tags per paper
#
tab_tags_breadth <- function(data, id_col, tags_col,
                             drop_vals = c("na", "indeterminable", "")) {
  data %>%
    select({{ id_col }}, {{ tags_col }}) %>%
    separate_rows({{ tags_col }}, sep = ";") %>%
    mutate({{ tags_col }} := str_trim({{ tags_col }})) %>%
    filter(!is.na({{ tags_col }}), !({{ tags_col }} %in% drop_vals)) %>%
    distinct({{ id_col }}, {{ tags_col }}) %>%
    count({{ id_col }}, name = "n_tags")
}

# ============================================================
# EXAMPLES (these are the lines you actually run per column)
# ============================================================

# ------------------------------------------------------------
# A) Factor-like columns (one value per row)
# ------------------------------------------------------------
tab_overarching_domain <- tab_factor(dat_std, overarching_domain)
tab_administration_mode <- tab_factor(dat_std, administration_mode)
tab_compound_items_codes <- tab_factor(dat_std, compound_items_codes)
tab_categorization_approach <- tab_factor(dat_std, categorization_approach)

tab_overarching_domain
tab_administration_mode
tab_compound_items_codes
tab_categorization_approach


# ------------------------------------------------------------
# B) Tag columns: DATABASE
# ------------------------------------------------------------
# 1) Which databases were used most often?
db_usage <- tab_tags_usage(dat_std, study_id, database)

# 2) How many databases per paper?
db_breadth <- tab_tags_breadth(dat_std, study_id, database)

# 3) Summaries for breadth
db_breadth_summary <- db_breadth %>%
  summarise(
    n_coded = n(),                 # how many studies had at least one database token
    mean = mean(n_tags),
    median = median(n_tags),
    min = min(n_tags),
    max = max(n_tags),
    n_single_db = sum(n_tags == 1) # how many used exactly 1 database
  )

db_usage
db_breadth_summary

# Distribution of number of databases (1,2,3,...)
db_breadth_dist <- db_breadth %>%
  count(n_tags, name = "n_papers") %>%
  mutate(percent = scales::percent(n_papers / sum(n_papers), accuracy = 0.1)) %>%
  arrange(n_tags)

db_breadth_dist


# ------------------------------------------------------------
# C) Tag columns: ITEM UNIT
# ------------------------------------------------------------
item_unit_usage <- tab_tags_usage(dat_std, study_id, item_unit)
item_unit_breadth <- tab_tags_breadth(dat_std, study_id, item_unit)

item_unit_usage

# how many papers have >1 unit token (proxy for "multiple analyses/units")
item_unit_multi <- item_unit_breadth %>%
  summarise(
    n_total = n(),
    n_multi = sum(n_tags > 1),
    percent_multi = scales::percent(n_multi / n_total, accuracy = 0.1)
  )

item_unit_multi


# ------------------------------------------------------------
# D) Tag columns: DIAGNOSTIC FRAMEWORK
# ------------------------------------------------------------
diag_fw_usage <- tab_tags_usage(dat_std, study_id, diagnostic_framework)
diag_fw_breadth <- tab_tags_breadth(dat_std, study_id, diagnostic_framework)

diag_fw_usage

diag_fw_breadth %>%
  summarise(
    n_coded = n(),
    mean = mean(n_tags),
    median = median(n_tags),
    min = min(n_tags),
    max = max(n_tags)
  )

```

```{r}


```