---
title: "02_methods_descriptives"
author: "Tim"
date: "2026-01-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(janitor)
library(rlang)
library(scales)# for percent formatting
library(here) # robust paths within the repo
```

```{r}
path_in <- here("data", "ica_methods.csv")

dir.create(here("outputs"), showWarnings = FALSE)

raw <- readr::read_csv(
file = path_in,
col_types = readr::cols(.default = readr::col_character()),
na = c("", "NA", "N/A", "na", "n/a", "null", "NULL")
)

dat <- raw %>% clean_names()

glimpse(dat) #sanity check
```

```{r}

binary_cols <- c(
  "matrix",
  "short_scales",
  "scale_search",
  "prisma_a",
  "prisma",
  "inclusion_criteria_applicable",
  "inclusion_binary",
  "clinical_relevance",
  "language_mentioned",
  "availability_mentioned",
  "release_date_mentioned",
  "frequency",
  "min_scale_citation",
  "exclusion_criteria",
  "within_scale_item_collapse",
  "compound_items",
  "idiosyncratic_items",
  "reverse_worded_items",
  "coding_team_reported",
  "blinded",
  "codebook",
  "software_mentioned",
  "item_exclusions",
  "irr_reported",
  "preregistered_rules",
  "coding_scheme_independent_mentioned",
  "mean_overlap",
  "min_overlap",
  "max_overlap",
  "length_correlation_noted"
)

# Select only the binary columns (keep a copy of IDs)
dat_binary_raw <- dat %>%
  select(study_id, any_of(binary_cols))
``` 

```{r}
to_logical_strict <- function(x, colname) {
  # Leave logical as-is
  if (is.logical(x)) return(x)

  x0 <- as.character(x)
  x1 <- tolower(trimws(x0))

  true_vals  <- c("true", "t", "yes", "y", "1")
  false_vals <- c("false", "f", "no", "n", "0")

  out <- dplyr::case_when(
    is.na(x1) ~ NA,
    x1 %in% true_vals  ~ TRUE,
    x1 %in% false_vals ~ FALSE,
    TRUE ~ NA
  )

  # Catch typos: any non-NA value not in the allowed encodings
  bad <- which(!is.na(x1) & !(x1 %in% c(true_vals, false_vals)))
  if (length(bad) > 0) {
    bad_vals <- sort(unique(x0[bad]))
    rlang::abort(paste0(
      "Non-boolean value(s) detected in column '", colname, "': ",
      paste(bad_vals, collapse = ", "),
      ". Example row indices: ", paste(head(bad, 10), collapse = ", ")
    ))
  }

  out
}

dat_binary <- dat_binary_raw %>%
  mutate(across(all_of(binary_cols), ~ to_logical_strict(.x, cur_column())))

```

```{r}
binary_counts <- dat_binary %>%
  pivot_longer(  #long table with one row per (paper × variable)
    cols = all_of(binary_cols),
    names_to = "variable",
    values_to = "value"
  ) %>%
  mutate(value = case_when( #converts logicals into `"TRUE"/"FALSE"/"NA"` labels
    is.na(value) ~ "NA",
    value ~ "TRUE",
    !value ~ "FALSE"
  )) %>%
  count(variable, value, name = "n") %>%  #counts how many of each label occurs per variable
  group_by(variable) %>%
  mutate(n_total = sum(n)) %>% #computes the total number of observations per variable while still in long format
  ungroup() %>%
  pivot_wider(  #spreads the counts into columns 'TRUE', 'FALSE' and 'NA' and keeps the n_total
    names_from = value,
    values_from = n,
    values_fill = 0
  ) %>%
  arrange(desc(TRUE))

binary_counts
```

```{r}
binary_overall <- dat_binary %>%
  summarise(  #computes per-column counts
    across(
      all_of(binary_cols),
      list(
        n_true  = ~ sum(.x %in% TRUE, na.rm = TRUE),
        n_false = ~ sum(.x %in% FALSE, na.rm = TRUE),
        n_na    = ~ sum(is.na(.x)),
        n_coded = ~ sum(!is.na(.x))
      ),
      .names = "{.col}__{.fn}"  #attempt to give systematic names like "irr_reported_n_true"
    )
  ) %>%
  pivot_longer(  #splits the names into 'variable" (the original column name) and '.value' (which statistic the number belongs to)
    cols = everything(),
    names_to = c("variable", ".value"),
    names_sep = "__"
  ) %>%
  mutate(
    prop_true = n_true / n_coded,  #proportion TRUE among all coded cases only
    percent_true = scales::percent(prop_true, accuracy = 0.1),
    report = paste0(n_true, "/", n_coded, " (", percent_true, ")")  #manuscript-friendly string
  ) %>%
  arrange(desc(prop_true))

binary_overall %>% select(variable, n_true, n_coded, n_na, report)

readr::write_csv(binary_overall, here("outputs", "binary_overall.csv"))

```

# Step 4: Conditional reporting rates for gated variables
There are a couple of key "gates" in the binary dataset: 
1. `prisma` only matters when `prisma_a == TRUE`
2. “selection criteria” only matter when `scale_search == TRUE`

This is done as: **filter to applicable subset → summarise the target variables**

## Step 4a: PRISMA reporting among applicable studies (`prisma_a == TRUE`)
### Why this step?
It enables to report a conditional statement and avoids penalising studies where PRISMA is structurally not applicable:

> “Among studies that conducted their own scale search (PRISMA applicable), X% reported a PRISMA-style flow diagram.”

```{r}
prisma_cond <- dat_binary %>%
  filter(prisma_a == TRUE) %>%
  summarise(
    n_applicable = n(),
    n_true  = sum(prisma %in% TRUE, na.rm = TRUE),
    n_false = sum(prisma %in% FALSE, na.rm = TRUE),
    n_na    = sum(is.na(prisma)),
    n_coded = sum(!is.na(prisma))
  ) %>%
  mutate(
    prop_true = n_true / n_coded,
    percent_true = scales::percent(prop_true, accuracy = 0.1),
    report = paste0(n_true, "/", n_coded, " (", percent_true, ")")
  )

prisma_cond

```


## Step 4b: Selection-criteria reporting among manual searches (`scale_search == TRUE`)

```{r}
criteria_cols <- c(
  "inclusion_binary",
  "clinical_relevance",
  "language_mentioned",
  "availability_mentioned",
  "release_date_mentioned",
  "frequency",
  "min_scale_citation",
  "exclusion_criteria"
)

criteria_cond <- dat_binary %>%
  filter(scale_search == TRUE) %>%
  summarise(
    across(
      all_of(criteria_cols),
      list(
        n_true  = ~ sum(.x %in% TRUE, na.rm = TRUE),
        n_false = ~ sum(.x %in% FALSE, na.rm = TRUE),
        n_na    = ~ sum(is.na(.x)),
        n_coded = ~ sum(!is.na(.x))
      ),
      .names = "{.col}__{.fn}"
    )
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = c("variable", ".value"),
    names_sep = "__"
  ) %>%
  mutate(
    prop_true = n_true / n_coded,
    percent_true = scales::percent(prop_true, accuracy = 0.1),
    report = paste0(n_true, "/", n_coded, " (", percent_true, ")"),
    condition = "Among studies with manual scale search (scale_search == TRUE)"
  ) %>%
  arrange(desc(prop_true))

criteria_cond %>% select(condition, variable, n_true, n_coded, n_na, report)

```

```{r}
readr::write_csv(prisma_cond, here("outputs", "prisma_conditional.csv"))
readr::write_csv(criteria_cond, here("outputs", "criteria_conditional.csv"))

```

#Descriptives for "homogenous" numeric columns


```{r}
# "clean"/homogenous numeric columns 
numeric_cols_clean <- c(
  "n_scales",
  "initial_coding_team_size",
  "n_rater",
  "mean_overlap_nr",
  "min_overlap_nr",
  "max_overlap_nr"
)

numeric_cols_clean_present <- intersect(numeric_cols_clean, names(dat))

dat_num <- dat %>%
  mutate(across(all_of(numeric_cols_clean_present), readr::parse_number))

# NA rates after parsing (quick QA)
numeric_na_rates <- dat_num %>%
  summarise(across(all_of(numeric_cols_clean_present),
                   ~ mean(is.na(.x)))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "prop_na") %>%
  mutate(percent_na = scales::percent(prop_na, accuracy = 0.1)) %>%
  arrange(desc(prop_na))

numeric_na_rates

``` 

```{r}
summarise_numeric <- function(x) {
  x <- x[!is.na(x)]
  if (length(x) == 0) {
    return(tibble(
      n = 0, mean = NA_real_, sd = NA_real_, median = NA_real_, iqr = NA_real_,
      min = NA_real_, max = NA_real_, q25 = NA_real_, q75 = NA_real_,
      n_outliers_iqr = NA_integer_
    ))
  }

  q <- quantile(x, probs = c(.25, .75), names = FALSE)
  iqr_val <- q[2] - q[1]
  lower <- q[1] - 1.5 * iqr_val
  upper <- q[2] + 1.5 * iqr_val
  n_out <- sum(x < lower | x > upper)

  tibble(
    n = length(x),
    mean = mean(x),
    sd = sd(x),
    median = median(x),
    iqr = IQR(x),
    min = min(x),
    max = max(x),
    q25 = q[1],
    q75 = q[2],
    n_outliers_iqr = n_out
  )
}

numeric_overall <- dat_num %>%
  summarise(across(all_of(numeric_cols_clean_present), summarise_numeric)) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "stats") %>%
  unnest_wider(stats)

numeric_overall

```


#Conditional descriptives for overlap values (only where Jaccard values are reported)
```{r}
overlap_pairs <- tribble(
  ~flag, ~value,
  "mean_overlap", "mean_overlap_nr",
  "min_overlap",  "min_overlap_nr",
  "max_overlap",  "max_overlap_nr"
) %>%
  filter(flag %in% names(dat_num), value %in% names(dat_num))

overlap_conditional <- overlap_pairs %>%
  pmap_dfr(function(flag, value) {

    # 1) subset to rows where the metric is reported/applicable
    df_sub <- dat_num %>% filter(.data[[flag]] == TRUE)

    # 2) pull the numeric vector from the relevant *_nr column
    x <- df_sub %>% pull(.data[[value]])

    # 3) compute descriptives on the numeric vector
    summarise_numeric(x) %>%
      mutate(
        flag = flag,
        value = value,
        n_applicable = nrow(df_sub)  # how many rows satisfied flag == TRUE
      ) %>%
      relocate(flag, value, n_applicable)
  })

overlap_conditional

```

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
